---
title: "CAS MA 214 — Project 2 Week 3"
subtitle: "Real Data Analysis"
author: "Group Members: ______________________________________"
date: "Wednesday, March 25, 2026"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

------------------------------------------------------------------------

## Group Setup

**Group members:**

| Name | BUID | Present/Absent |
|------|------|----------------|
|      |      |                |
|      |      |                |
|      |      |                |
|      |      |                |

**Dataset:** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Response variable (**$Y$): \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Primary predictor (**$X$): \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

------------------------------------------------------------------------

# Overview

Last week you built and ran a simulation study where the true $\beta_1$
was **known**. This week you apply the same tools — OLS, bootstrap CI,
and bootstrap p-value — to **real data**, where the truth is unknown.

Your goals for this lab are:

1.  Load and explore your dataset.
2.  Fit OLS and compute the classical CI and p-value for the slope.
3.  Compute the bootstrap CI and bootstrap p-value.
4.  Run residual diagnostics to check OLS assumptions.
5.  Connect what you see in the real data back to your simulation
    results from last week.

------------------------------------------------------------------------

# Part 1: Load and Explore Your Data

## Step 1: Load the Dataset

Load your group's chosen dataset and make sure it is clean and ready to
use. Your data frame should be called `dat` throughout this lab.

```{r load-data}
# TODO: Load your dataset.

# Check the result
head(dat)
dim(dat)
```

```{r summary-stats}
# Summary statistics for Y and X
summary(dat[, c("Y", "X")])


# TODO: Drop rows with missing Y or X if needed (optional)

```

#### Discussion Question 1. Describe your dataset in 2–3 sentences. What does each row represent? What are the units of $Y$ and $X$?

-\>

------------------------------------------------------------------------

## Step 2: Exploratory Data Analysis (EDA)

```{r eda-plot}
# TODO: Scatterplot of Y vs X with OLS line

```

#### Discussion Question 2. Does the relationship look roughly linear? Note any curvature, fanning (heteroskedasticity), or outliers. Which of your two simulation studies does this scatterplot most remind you of?

-\>

------------------------------------------------------------------------

# Part 2: OLS Fit and Classical Inference

## Step 3: Fit the Model and Extract the Classical CI

```{r ols-fit}
alpha <- 0.05

fit <- lm(Y ~ X, data = dat)
s   <- summary(fit)
s
```

```{r classical-ci}
# TODO: Extract slope, SE, df, and compute the CI manually
# (same approach as last week's simulation)


```

#### Discussion Question 3. Interpret the slope $\hat\beta_1$ in context. What does a one-unit increase in $X$ correspond to in terms of $Y$? Is the classical p-value significant at the 5% level?

-\>

------------------------------------------------------------------------

# Part 3: Bootstrap CI and Bootstrap P-value

## Step 4: Run the Bootstrap

This follows the same structure as Step 2 from last week, but now
applied to your real dataset instead of simulated data.

```{r bootstrap}

B <- 1000   # number of bootstrap resamples
n <- nrow(dat)

boot_b1 <- numeric(B)

for (j in seq_len(B)) {
  idx        <- sample(n, replace = TRUE)   # resample row        # TODO: save the slope
}

# Bootstrap CI (percentile method)
ci_boot <- quantile(boot_b1, c(0.025, 0.975))

cat("Classical CI: [", round(ci_lower,    3), ",", round(ci_upper,    3), "]\n")
cat("Bootstrap CI: [", round(ci_boot[1],  3), ",", round(ci_boot[2],  3), "]\n")
```

```{r bootstrap-pvalue}
# Bootstrap p-value for H0: beta1 = 0
# Shift the bootstrap distribution to be centred at 0 (impose the null),
# then find the proportion of resamples as extreme as the observed b1.

# TODO: pval_boot

cat("Classical p-value: ", round(pval_cls,  4), "\n")
cat("Bootstrap p-value: ", round(pval_boot, 4), "\n")
```

## Step 5: Plot the Bootstrap Distribution

```{r bootstrap-plot}
# TODO : Plot bootstrap distribution to answer the discussion question 4 here. 
```

#### Discussion Question 4. Is the bootstrap distribution roughly symmetric and bell-shaped, or is it skewed? Do the classical and bootstrap CIs agree? Do the two p-values agree? If they differ, which would you trust more and why?

-\>

------------------------------------------------------------------------

# Part 4: Residual Diagnostics

## Step 6: Check OLS Assumptions

```{r diagnostics}
par(mfrow = c(1, 2))

# TODO: Residuals vs Fitted — looks for nonlinearity and heteroskedasticity


# TODO: Normal Q-Q — checks normality of residuals

```

```{r residual-histogram}
# TODO: Any more plots to check or show?
```

#### Discussion Question 5. Describe what you see in the residual plots.

Fill in the table:

| Diagnostic plot | What you see | Assumption it checks | Violated? (Yes / No / Unclear) |
|------------------|------------------|------------------|------------------|
| Residuals vs Fitted |  | Linearity |  |
| Residuals vs Fitted |  | Constant variance |  |
| Normal Q-Q |  | Normality of errors |  |
| Residual histogram |  | Normality of errors |  |

-\>

------------------------------------------------------------------------

# Part 5: Connect Back to Your Simulation

## Step 7: Which Simulation Study Does This Resemble?

#### Discussion Question 6. Look at the violation pattern you identified in Step 6. Which of your two simulation studies (from last week) does your real data most resemble? Fill in the table.

|                     | Your real data | Most similar simulation scenario |
|---------------------|----------------|----------------------------------|
| Linearity           |                |                                  |
| Constant variance   |                |                                  |
| Normality of errors |                |                                  |
| Sample size $n$     |                |                                  |

-\>

#### Discussion Question 8. Did the bootstrap CI and bootstrap p-value give meaningfully different answers from the classical ones? Given what you found in the diagnostics, is that what you expected?

-\>

------------------------------------------------------------------------

## Step 8: Plan for the Report

#### Discussion Question 9. What is the main conclusion your group will draw in Part II of the report? Write 2–3 sentences summarising: what violation you detected, which simulation study it matches, and what that implies about trust in the classical CI.

-\>

------------------------------------------------------------------------

*Submit this `.Rmd` and its knitted PDF to Blackboard **at the end of
the lab today**.*

## Reminder : Deliverable 2 (Progress Report; Slides, R code and the data) due Monday March 30th 10pm.
