---
title: "CAS MA 214 — Project 2 Lab 1"
subtitle: "Bootstrapping, Data Exploration, and Study Selection"
author: "Group Members: ______________________________________"
date: "Wednesday, March 4, 2026"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

------------------------------------------------------------------------

## Group Setup

**Group members:**

1.  

    ------------------------------------------------------------------------

2.  

    ------------------------------------------------------------------------

3.  

    ------------------------------------------------------------------------

4.  

    ------------------------------------------------------------------------

------------------------------------------------------------------------

# Part 1: What Is Bootstrapping?

## The Problem with Classical CIs

The classical confidence interval for a regression slope uses the
formula:

$$\hat{\beta}_1 \pm t^* \cdot \widehat{SE}(\hat{\beta}_1)$$

The standard error $\widehat{SE}$ is derived under specific assumptions:
normally distributed errors and constant variance. When those
assumptions fail, the formula may be wrong --- the CI may be too narrow
or too wide, and may no longer contain the true slope 95% of the time.

The **bootstrap** is a different approach. Instead of relying on a
formula, it uses the data itself to approximate the sampling variability
of $\hat{\beta}_1$ --- without assuming anything about the error
distribution.

## The Core Idea

The bootstrap answers the question: *if I could draw many new samples
from the same population, how much would* $\hat{\beta}_1$ *vary?*

Since we can't actually draw new samples from the population, we
approximate it by treating our dataset as the population and sampling
from it **with replacement**.

**One bootstrap iteration:**

1.  Take your dataset of $n$ rows.
2.  Sample $n$ rows **with replacement** --- some rows appear twice or
    more, some not at all.
3.  Fit OLS on this *bootstrap sample*. Record $\hat{\beta}_1^*$.

Repeat $B$ times. The resulting $B$ slope estimates form the **bootstrap
distribution** --- an approximation of the true sampling distribution of
$\hat{\beta}_1$. The middle 95% of that distribution is the **bootstrap
percentile CI**.

------------------------------------------------------------------------

## Step 1.1: A Toy Example --- One Resample by Hand

We start with a small, made-up dataset so the mechanics are transparent.

```{r toy-data}
set.seed(214)

# A simple toy dataset: 10 observations
n <- 10
x <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
e <- NULL # TODO : e is error term; make it
y <- 2 + 1.5 * x #+ e

toy_data <- data.frame(x = x, y = y)
toy_data

# Fit OLS on the original data and record the slope
 fit_toy    <- lm(y ~ x, data = toy_data)
 slope_orig <- coef(fit_toy)["x"]
 cat("Original slope:", round(slope_orig, 3), "\n")
```

Now draw one bootstrap sample and refit.

```{r one-resample}

# Sample n row indices WITH replacement
# Hint: sample(nrow(toy_data), replace = TRUE)
index <- NULL # TODO: sample index from toy_data with replacement

# Show which rows were selected (some will appear more than once)
cat("Selected rows:", index, "\n")
toy_data[index, ]
```

```{r one-boot-slope}
# Fit Simple Linear Regression on the bootstrap sample and extract the slope
# Hint: lm(), coef()
fit_boot   <- NULL # TODO
slope_boot <- NULL # TODO

cat("Original slope:", round(slope_orig, 3), "\n")
cat("Bootstrap slope:", round(slope_boot, 3), "\n")
```

**Q1.** Is the bootstrap slope the same as the original? Why or why not?
What effect does sampling with replacement have on the fit?

> *Answer here.*

------------------------------------------------------------------------

## Step 1.2: Repeat $B$ Times

Now repeat the resample--refit--record process $B = 500$ times.

```{r bootstrap-loop-toy}
B_toy        <- 500
boot_slopes  <- numeric(B_toy)   # pre-allocate storage

for (j in seq_len(B_toy)) {
  index            <- NULL #TODO   # sample row indices with replacement
  fit_b          <- NULL #TODO   # refit lm on toy_data[index, ]
  boot_slopes[j] <- NULL #TODO   # store the slope  -- hint: coef()["x"]
}

cat("First 10 bootstrap slopes:", round(boot_slopes[1:10], 3), "\n")
```

------------------------------------------------------------------------

## Step 1.3: Plot the Bootstrap Distribution

```{r boot-dist-toy}
# Plot a histogram of boot_slopes
# Add a vertical line at slope_orig (original slope)
# Add vertical lines at the 2.5th and 97.5th percentiles
# Hint: ggplot() + geom_histogram() + geom_vline()

# TODO 

```

**Q2.** Describe the shape of the bootstrap distribution. Where is it
centred? What does the spread of this distribution represent?

> *Answer here.*

------------------------------------------------------------------------

## Step 1.4: Compute the Bootstrap CI

```{r boot-ci-toy}
# 95% bootstrap percentile CI: 2.5th and 97.5th percentiles
# Hint: quantile()
ci_boot <- NULL #TODO
cat("Bootstrap 95% CI: [", round(ci_boot[1], 3),
    ",", round(ci_boot[2], 3), "]\n")

# Classical 95% CI from the original fit
ci_classic <- NULL #TODO
```

**Q3.** The true slope in this toy example is $\beta_1 = 1.5$. Do both
CIs contain it? Are the two CIs similar in width and location, or do
they differ noticeably? Why might they agree or disagree here?

> *Answer here.*

------------------------------------------------------------------------

# Part 2: Apply Bootstrap to Your Own Dataset

Now apply what you just built to your group's real dataset for Part II
of the project.

## Step 2.1: Choose a Dataset

Your dataset must have a **continuous response variable**, at least one
**continuous predictor**, and at least **50 observations**.

Possible sources (or propose your own):

-   `Boston` in the `MASS` package --- median home value vs.
    socioeconomic predictors
-   `gapminder` in the `gapminder` package --- life expectancy vs. GDP
    per capita
-   Auto MPG --- [UCI ML Repository](https://archive.ics.uci.edu/)
-   [OpenIntro Datasets](https://www.openintro.org/data/)
-   [Kaggle](https://www.kaggle.com/datasets)
-   [data.gov](https://data.gov/)

**Our dataset:** \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Response variable (**$Y$):
\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Primary predictor (**$X$):
\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

**Why this pair?** (1--2 sentences)

> *Answer here.*

------------------------------------------------------------------------

## Step 2.2: Load, Summarise, and Plot

```{r load-data}
# Load your dataset
# Hint: library(), data(), read.csv(), or similar


# Summary statistics for your response and predictor
# Hint: summary()

```

```{r eda-plot}
# Scatterplot with linear regression line
# Hint: ggplot() + geom_point() + geom_smooth(method = "lm")

```

**Q4.** Does the relationship look roughly linear? Describe direction,
strength, any curvature, outliers, or fanning of the points.

> *Answer here.*

------------------------------------------------------------------------

## Step 2.3: Fit OLS and Run Diagnostics

```{r ols-fit}
# Fit lm, print summary, and compute classical 95% CI for the slope
# Hint: lm(), summary()

```

```{r diagnostics}
# Residuals vs Fitted  -- hint: plot(fit, which = 1)
# Normal Q-Q           -- hint: plot(fit, which = 2)

```

**Q5.** Do the residuals look roughly normal and evenly spread? Note any
patterns --- curvature, fanning, or heavy tails.

> *Answer here.*

------------------------------------------------------------------------

## Step 2.4: Bootstrap CI on Your Data

Adapt your loop from Step 1.2 to your own dataset. Use $B = 100$.

```{r bootstrap-real}

B    <- 1000
n    <- nrow(dat)   # make sure your data frame is called dat, or adjust

boot_slopes_real <- numeric(B)

for (j in seq_len(B)) {
  index                  <- NULL # TODO
  fit_b                 <- NULL # TODO
  boot_slopes_real[j]    <- NULL # TODO
}

# Bootstrap CI
ci_boot_real   <- NULL # TODO

# Print both CIs side by side
cat("Classical CI: [", round(ci_classical_real[1], 3),
    ",", round(ci_classical_real[2], 3), "]\n")
cat("Bootstrap CI: [", round(ci_boot_real[1],     3),
    ",", round(ci_boot_real[2],     3), "]\n")
```

```{r boot-dist-real}
# Plot the bootstrap distribution for your real data
# Mark the original slope and the CI bounds
# Hint: ggplot() + geom_histogram() + geom_vline()

```

**Q6.** Are the two CIs similar or different? Is the bootstrap
distribution symmetric? Given your diagnostic plots in Step 2.3, which
CI would you trust more and why?

> *Answer here.*

------------------------------------------------------------------------

# Part 3: Study Selection

In Part I of the project you will run a controlled simulation where
$\beta_1 = 2$ is known, and directly measure how often each CI method
captures it under different assumption violations.

## Step 3.1: The Four Studies

All groups use the same base model:

$$Y_i = \beta_0 + \beta_1 X_i + \varepsilon_i, \qquad \beta_0 = 1,\ \beta_1 = 2,\ X_i \sim N(0,1)$$

The only thing that varies across studies is the distribution of
$\varepsilon_i$:

| Study | Violation | What changes across scenarios |
|----|----|----|
| A | Non-normality | Skewness of the error distribution |
| B | Heteroskedasticity | Error variance grows with $\|X\|$ |
| C | Heavy-tailed errors | Degrees of freedom $\nu$ of a $t$-error |
| D | Nonlinearity | Quadratic coefficient in the true DGM (Data Generating Model) |

Each study has four scenarios: a **baseline** (normal errors, no
violation) plus three increasing severity levels.

------------------------------------------------------------------------

## Step 3.2: Choose Your Two Studies

**Our two chosen studies:** Study \_\_\_\_\_ and Study \_\_\_\_\_

**Why Study 1?** (1--2 sentences)

> *Answer here.*

**Why Study 2?** (1--2 sentences)

> *Answer here.*

**Connection to your real data:** Does either chosen study match what
you saw in the diagnostic plots from Step 2.3? Explain briefly.

> *Answer here.*

------------------------------------------------------------------------

## Step 3.3: Predict Your Results

**Q7.** For each chosen study: as the violation gets more severe, do you
expect the **classical CI coverage** to drop below 0.95, stay roughly at
0.95, or vary unpredictably? What about the **bootstrap CI**?

> *Study 1 prediction:*

> *Study 2 prediction:*

**Q8.** The baseline scenario (normal errors, $n = 30$) is the same for
all groups. What coverage do you expect for both CIs there, and why?
What would it mean if you ran the baseline and got coverage of 0.75?

> *Answer here.*

------------------------------------------------------------------------

## Wrap-Up

**Before you leave, confirm your group has:**

-   [ ] Bootstrap loop built and working on the toy example (Steps
    1.2--1.4)
-   [ ] Dataset chosen, loaded, and EDA plot produced (Steps 2.1--2.2)
-   [ ] OLS fit, diagnostics, and bootstrap CI done on real data (Steps
    2.3--2.4)
-   [ ] Two studies chosen and justified (Step 3.2)
-   [ ] All eight discussion questions answered

------------------------------------------------------------------------

*Submit this `.Rmd` and its knitted PDF to Blackboard by **End of the
Lab today**. One submission per group.*

------------------------------------------------------------------------

**Post-Lab Activity**

You have Deliverable 1 to submit - Project Outline due on Monday 16 at
10pm.
