---
title: "Lab 05: The Coffee Shop Oracle"
subtitle: "MA214 Applied Statistics — Bayesian Linear Regression"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.width  = 10,
  fig.height = 6,
  fig.align  = "center",
  out.width  = "95%"
)
library(tidyverse)   # ggplot2, dplyr, purrr, tidyr
library(broom)       # tidy model outputs
library(patchwork)   # combining ggplots
library(coda)        # effectiveSize()
```

```{r helpers, include=FALSE}
# ── Posterior summary ─────────────────────────────────────────────────
# Takes a full MCMC chain, discards the first `burn` iterations,
# and returns a one-row data frame with key posterior diagnostics.
posterior_summary <- function(chain, burn = 1000, prob = 0.95) {
  s  <- chain[-(1:burn)]                                   # discard burn-in
  ci <- quantile(s, c((1 - prob) / 2, 1 - (1 - prob) / 2)) # equal-tail CrI
  ci <- as.vector(ci)
  data.frame(
    Mean        = round(mean(s), 3),
    SD          = round(sd(s),   3),
    `CrI Lower` = round(ci[1],  3),
    `CrI Upper` = round(ci[2],  3),
    check.names = FALSE
  )
}

# ── Shared theme ──────────────────────────────────────────────────────
lab_theme <- function(base_size = 13) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      plot.title       = element_text(face = "bold", hjust = 0.5,
                                      size = base_size + 3),
      plot.subtitle    = element_text(hjust = 0.5, size = base_size - 1,
                                      colour = "grey40"),
      panel.grid.minor = element_blank(),
      strip.text       = element_text(face = "bold"),
      legend.position  = "bottom"
    )
}
```

------------------------------------------------------------------------

## The Scenario

**BU Coffee Shop** wants to predict **daily revenue** from the number of
customers. We model the relationship as:

$$y_i = \beta_0 + \beta_1\,x_i + \varepsilon_i,
\qquad \varepsilon_i \overset{iid}{\sim} N(0,\,\sigma^2)$$

| Symbol     | Meaning                        |
|------------|--------------------------------|
| $y_i$      | Daily revenue (\$)             |
| $x_i$      | Customers per day              |
| $\beta_0$  | Baseline revenue (0 customers) |
| $\beta_1$  | Marginal revenue per customer  |
| $\sigma^2$ | Unexplained variance           |

------------------------------------------------------------------------

# Step 1: Simulate Data

We generate synthetic data from a known truth so we can check later
whether our Bayesian sampler recovers the correct parameter values.

**Key ideas:**

- `x ~ Uniform(0, 10)` plays the role of the number of customers on a
  given day.
- The noise `rnorm(n, 0, sigma_true)` adds realistic measurement scatter
  around the true line.
- We store everything in a `data.frame` called `coffee` for easy
  reference later.

```{r step1-simulate}
set.seed(2026214)

n          <- 100        # number of observed days
x          <- runif(n, 0, 10)    # predictor: customers per day
beta0_true <- 5         # true intercept ($5 baseline)
beta1_true <- 2         # true slope ($2 per customer)
sigma_true <- 1         # true residual SD

y      <- beta0_true + beta1_true * x + rnorm(n, 0, sigma_true)
coffee <- data.frame(revenue = y, customers = x)

ggplot(coffee, aes(x = customers, y = revenue)) +
  geom_point(colour = "#1f77b4", size = 2.5, alpha = 0.75) +
  geom_smooth(method = "lm", se = TRUE,
              colour = "#d62728", fill = "#d62728",
              alpha = 0.15, linewidth = 1.1) +
  geom_abline(intercept = beta0_true, slope = beta1_true,
              colour = "#2ca02c", linewidth = 1, linetype = "dashed") +
  annotate("text",
           x = 0.3, y = beta0_true + beta1_true * 0.3 + 1.5,
           label = "True line", colour = "#2ca02c",
           size = 4, hjust = 0) +
  labs(
    title    = "Revenue vs Customers \u2014 Simulated Data",
    subtitle = paste0("True: beta0 = ", beta0_true,
                      ",  beta1 = ", beta1_true,
                      ",  sigma = ", sigma_true),
    x = "Customers per day",
    y = "Revenue ($)"
  ) +
  lab_theme()
```

------------------------------------------------------------------------

# Step 2: Bayesian Model

## 2.1 Specification

The Bayesian model has three layers:

**Likelihood** — how the data arise given parameters:
$$y_i \mid \beta_0, \beta_1, \sigma^2
  \;\sim\; N\!\left(\beta_0 + \beta_1 x_i,\; \sigma^2\right)$$

**Priors** — our beliefs about the parameters *before* seeing data:
$$\beta_0 \sim N(5,\,2^2), \qquad
\beta_1 \sim N(2,\,0.5^2), \qquad
\sigma^2 \sim \text{Log-Normal}(\log 1,\,0.3^2)$$

> **Informative vs. weakly informative priors — why it matters here:**
>
> In the original (weakly informative) version we used
> $\beta_1 \sim N(0, 10^2)$ and $\sigma^2 \sim \text{LogNormal}(0, 1)$.
> Both priors placed substantial probability mass over implausible
> values (e.g., negative revenue-per-customer, or residual variance of
> 100). Here we use **informative priors** that encode genuine domain
> knowledge about a coffee shop:
>
> - $\beta_0 \sim N(5, 2^2)$: Baseline revenue near \$5 with SD \$2.
>   Still moderately uncertain, but rules out values like −\$20 or
>   \$100.
> - $\beta_1 \sim N(2, 0.5^2)$: We believe each additional customer
>   brings in roughly \$2, with a tight SD of \$0.50. For example, a
>   prior 95% interval is approximately $2 \pm 1.96 \times 0.5 =
>   [1.02,\; 2.98]$ — sensible for a café menu where average spend is a
>   few dollars.
> - $\sigma^2 \sim \text{LogNormal}(\log 1, 0.3^2)$: Prior median is
>   $e^{\log 1} = 1$, and the tighter SD of 0.3 on the log scale gives a
>   95% interval of roughly $[e^{\log 1 - 1.96 \times 0.3},\;
>   e^{\log 1 + 1.96 \times 0.3}] \approx [0.55,\; 1.80]$ — we are
>   fairly confident day-to-day revenue noise is within about \$1–\$2.
>
> **Effect on the posterior:** With $n = 100$ observations the
> likelihood is dominant and both prior choices will give similar
> posteriors. The benefit of informative priors shows most clearly when
> (a) $n$ is small, (b) the chain needs fewer burn-in iterations to
> reach the posterior mode (it starts closer), and (c) the posterior is
> narrower and better-estimated.

**Posterior** (by Bayes' theorem, up to a normalizing constant):
$$\log p(\theta \mid y)
  \propto \underbrace{\log p(y \mid \theta)}_{\text{log-likelihood}}
  + \underbrace{\log p(\theta)}_{\text{log-prior}}$$

We work in **log space** because:

1.  Likelihoods are products of many small probabilities; on the natural
    scale these underflow to zero. Logs turn products into sums, keeping
    everything numerically stable.
2.  The normalizing constant $p(y)$ is just an additive constant in log
    space and cancels out in the MH acceptance ratio (Step 3).

## 2.2 Log-Posterior Function

The `log_posterior()` function below evaluates $\log p(\theta \mid y)$
at any candidate parameter triple $(\beta_0, \beta_1, \sigma^2)$.

**Breaking it down:**

| Line | What it computes | Why |
|------------------------|------------------------|------------------------|
| Guard | Returns $-\infty$ if $\sigma^2 \le 0$ | Variance must be strictly positive; $-\infty$ in log space means probability 0, causing MH to always reject. |
| `mu` | $\mu_i = \beta_0 + \beta_1 x_i$ for all $i$ | Vectorized mean prediction for each observation. |
| `log_lik` | $\sum_i \log N(y_i; \mu_i, \sigma^2)$ | Sum of log normal densities — the log-likelihood. |
| `log_p_beta0` | $\log N(\beta_0; 5, 4)$ | Log-prior for the intercept. |
| `log_p_beta1` | $\log N(\beta_1; 2, 0.25)$ | Log-prior for the slope — informative: \$2 ± \$0.50 per customer. |
| `log_p_sigma2` | $\log \text{LogNormal}(\sigma^2;\, \log 1, 0.3^2)$ | Log-prior for the variance — tighter than before; median 1, 95% range ≈ [0.55, 1.80]. |
| Return | Sum of all six terms | Log-posterior up to a constant. |

> **Why `dlnorm` for the** $\sigma^2$ prior? The Log-Normal distribution
> is a natural prior for strictly positive quantities: if
> $\sigma^2 \sim \text{LogNormal}(\mu_0, s_0^2)$, then
> $\log\sigma^2 \sim N(\mu_0, s_0^2)$. Here $\mu_0 = \log 1 = 0$ sets
> the prior median to exactly 1, and $s_0 = 0.3$ tightens the prior
> substantially relative to the weakly informative $s_0 = 1$: the 95%
> prior interval shrinks from $[0.14, 7.1]$ to approximately
> $[0.55, 1.80]$. This encodes the belief that day-to-day revenue
> scatter is on the order of \$1, not \$7. The code is simply
> `dlnorm(sigma2, meanlog = 0, sdlog = 0.3, log = TRUE)`.

```{r step2-logpost}
log_posterior <- function(beta0, beta1, sigma2) {

  # Guard: variance must be strictly positive
  # Returning -Inf means this point has probability 0 —
  # MH will always reject a proposal with sigma2 <= 0.
  if (sigma2 <= 0) return(-Inf)

  # ── Log-likelihood ────────────────────────────────────────────────
  # mu is the vector of fitted values for all 100 observations.
  # dnorm(..., log = TRUE) returns log densities; sum() gives the
  # total log-likelihood (i.e., log of the product of 100 densities).
  mu      <- beta0 + beta1 * x
  log_lik <- sum(dnorm(y, mean = mu, sd = sqrt(sigma2), log = TRUE))

  # ── Log-priors ────────────────────────────────────────────────────
  # Each dnorm / dlnorm call returns the log density of ONE parameter
  # under its prior distribution. We add these to the log-likelihood
  # because log(AB) = log(A) + log(B).

  log_p_beta0  <- dnorm(beta0,  mean = 5,  sd = 2,   log = TRUE)        # N(5, 2^2)
  log_p_beta1  <- dnorm(beta1,  mean = 2,  sd = 0.5, log = TRUE)        # N(2, 0.5^2) — informative
  log_p_sigma2 <- dlnorm(sigma2, meanlog = 0, sdlog = 0.3, log = TRUE)  # LogNormal(log1, 0.3^2)

  # ── Return log-posterior ──────────────────────────────────────────
  return_value = log_lik + log_p_beta0 + log_p_beta1 + log_p_sigma2
  return(return_value)
}

# ── Sanity checks ─────────────────────────────────────────────────────
# At the true parameter values the log-posterior should be finite.
# At sigma2 = -1 (impossible) it should return -Inf.
stopifnot(is.finite(log_posterior(5, 2, 1)))
stopifnot(log_posterior(5, 2, -1) == -Inf)
cat("log p(true params | y) =", round(log_posterior(5, 2, 1), 2), "\n")
```

------------------------------------------------------------------------

# Step 3: Metropolis–Hastings MCMC

## How MH Works

Markov chain Monte Carlo (MCMC) is a family of algorithms that **draw
samples from a target distribution** (here, the posterior) without
needing to know its normalizing constant. The **Metropolis–Hastings
(MH)** variant works as follows:

1.  Start at some initial parameter value $\theta^{(1)}$.
2.  At step $t$, **propose** a new value $\theta'$ from a proposal
    distribution $q(\theta' \mid \theta^{(t)})$.
3.  Compute the **log acceptance ratio**:
    $$\log \alpha = \log p(\theta' \mid y) - \log p(\theta^{(t)} \mid y)$$
4.  **Accept** $\theta^{(t+1)} = \theta'$ with probability
    $\min(1, e^{\log\alpha})$; otherwise keep
    $\theta^{(t+1)} = \theta^{(t)}$.
5.  Repeat.

> **Why log space for** $\alpha$? Computing
> $e^{\log p(\theta')} / e^{\log p(\theta)}$ directly can produce
> underflow/overflow. In log space, the ratio becomes a difference, and
> we compare $\log(\text{Uniform}(0,1))$ to $\log\alpha$ — both safely
> finite.

> **Why propose** $\sigma^2$ on the log scale? $\sigma^2$ is constrained
> to be positive. If we proposed on the natural scale with a symmetric
> normal, we could easily draw a negative value and always get rejected.
> Proposing $\log\sigma^2 \sim N(\log\sigma^2_\text{cur}, \tau^2)$
> guarantees positivity and explores the right-skewed support of
> $\sigma^2$ more efficiently. Note: this is an **asymmetric** proposal,
> but for the random-walk version used here the Jacobian cancels in the
> acceptance ratio.

**Acceptance rate tuning:**

| Rate   | Interpretation                              |
|--------|---------------------------------------------|
| \< 10% | Proposals too large — chain gets stuck      |
| 25–45% | Healthy — good mixing                       |
| \> 70% | Proposals too small — chain explores slowly |

Tune `prop_sd` to hit the 25–45% target.

```{r step3-mcmc}
set.seed(2006214)

n_iter <- 5000

# ── Burn-in ────────────────

burn <- 500   # initial placeholder;
# ── Storage vectors ───────────────────────────────────────────────────
# Pre-allocating with numeric(n_iter) is much faster than growing vectors.
beta0_chain  <- numeric(n_iter)
beta1_chain  <- numeric(n_iter)
sigma2_chain <- numeric(n_iter)

# ── Starting values ───────────────────────────────────────────────────
# We initialize at the informative prior means. Because our priors are
# centered close to the truth, the chain reaches the posterior mode
# quickly — this is one concrete benefit of informative priors: fewer
# burn-in iterations are needed.
beta0_chain[1]  <- 5    # prior mean of beta0
beta1_chain[1]  <- 2    # prior mean of beta1 (informative)
sigma2_chain[1] <- 1    # prior median of sigma2

# ── Proposal standard deviations ─────────────────────────────────────
# These control how far we jump each iteration.
# Smaller → higher acceptance, slower mixing.
# Larger  → lower acceptance (many rejections), also slower mixing.
# The values below were chosen to yield ~30% acceptance (see output).
prop_sd <- c(beta0 = 0.15, beta1 = 0.10, log_sigma2 = 0.20)

accept_count <- 0L

# ── Main MCMC loop ────────────────────────────────────────────────────
for (i in seq(2, n_iter)) {

  # Current state
  b0_cur <- beta0_chain[i - 1]
  b1_cur <- beta1_chain[i - 1]
  s2_cur <- sigma2_chain[i - 1]

  # Propose new values.
  # beta0 and beta1: symmetric normal random walk on natural scale.
  # sigma2: random walk on log scale, then exponentiate to ensure > 0.
  b0_prop <- rnorm(1, b0_cur, prop_sd["beta0"])
  b1_prop <- rnorm(1, b1_cur, prop_sd["beta1"])
  s2_prop <- exp(rnorm(1, log(s2_cur), prop_sd["log_sigma2"]))

  # Log acceptance ratio: log p(proposed) - log p(current)
  # If proposed is better (higher posterior), log_alpha > 0 → always accept.
  # If proposed is worse, log_alpha < 0 → accept with probability e^log_alpha.
  log_alpha <- log_posterior(b0_prop, b1_prop, s2_prop) -
               log_posterior(b0_cur,  b1_cur,  s2_cur)

  # Accept / reject
  # log(runif(1)) is log of U(0,1); comparing to log_alpha is equivalent
  # to comparing runif(1) < exp(log_alpha) but avoids overflow.
  if (log(runif(1)) < log_alpha) {
    beta0_chain[i]  <- b0_prop
    beta1_chain[i]  <- b1_prop
    sigma2_chain[i] <- s2_prop
    accept_count    <- accept_count + 1L
  } else {
    # Rejection: stay at current state (duplicate the previous row)
    beta0_chain[i]  <- b0_cur
    beta1_chain[i]  <- b1_cur
    sigma2_chain[i] <- s2_cur
  }
}

accept_pct <- accept_count / (n_iter - 1) * 100
cat(sprintf("Acceptance rate: %.1f%%  %s\n", accept_pct,
            ifelse(accept_pct > 20 & accept_pct < 50,
                   "\u2713 healthy", "\u26a0 consider tuning prop_sd")))
```

------------------------------------------------------------------------

# Step 4: Convergence Diagnostics

Before interpreting posterior summaries, we must verify that the chain
has **converged** to the posterior and is **mixing well**.

## 4.1 Trace Plots and Posterior Densities

**What to look for in trace plots:**

| Pattern | Meaning | Action |
|------------------------|------------------------|------------------------|
| "Hairy caterpillar" — noisy horizontal band | ✅ Well-mixed | None |
| Visible trend or drift | ❌ Not yet converged | Increase burn-in |
| Long horizontal runs (gets "stuck") | ❌ Poor mixing | Decrease `prop_sd` |
| Sudden jumps between two levels | ❌ Bimodal posterior or bad proposal | Investigate |

```{r step4-diagnostics, fig.height=9}
# post_burn is the first iteration index in the post-burn-in phase.
post_burn <- burn + 1

# ── Tidy chains into long format for ggplot ────────────────────────────
mcmc_df <- tibble(
  iter   = 1:n_iter,
  beta0  = beta0_chain,
  beta1  = beta1_chain,
  sigma2 = sigma2_chain
) |>
  pivot_longer(-iter, names_to = "param", values_to = "value") |>
  mutate(
    phase = if_else(iter <= burn, "Burn-in", "Post-burn-in"),
    param = factor(param,
                   levels = c("beta0", "beta1", "sigma2"),
                   labels = c("beta[0]", "beta[1]", "sigma^2"))
  )

# ── Trace plots ────────────────────────────────────────────────────────
p_trace <- ggplot(mcmc_df, aes(iter, value, colour = phase)) +
  geom_line(linewidth = 0.35, alpha = 0.9) +
  facet_wrap(~ param, scales = "free_y", ncol = 1,
             labeller = label_parsed) +
  scale_colour_manual(
    values = c("Burn-in" = "#e05c5c", "Post-burn-in" = "#3a7abf")
  ) +
  labs(title = "Trace Plots",
       subtitle = paste0("Red = burn-in (", burn, " iter, discarded)  |  Blue = post-burn-in (kept)"),
       x = "Iteration", y = "Value", colour = NULL) +
  lab_theme()

# ── True-value reference lines ─────────────────────────────────────────
true_vals <- tibble(
  param = factor(c("beta[0]", "beta[1]", "sigma^2"),
                 levels = c("beta[0]", "beta[1]", "sigma^2")),
  truth = c(beta0_true, beta1_true, sigma_true^2)
)

# ── Prior vs Posterior overlay ─────────────────────────────────────────
# For each parameter we draw the prior density curve alongside the
# posterior histogram so students can see how the data updated our beliefs.
prior_curves <- bind_rows(
  tibble(
    param = "beta[0]",
    x     = seq(0, 10, length.out = 300),
    dens  = dnorm(seq(0, 10, length.out = 300), mean = 5, sd = 2)
  ),
  tibble(
    param = "beta[1]",
    x     = seq(0.5, 3.5, length.out = 300),
    dens  = dnorm(seq(0.5, 3.5, length.out = 300), mean = 2, sd = 0.5)
  ),
  tibble(
    param = "sigma^2",
    x     = seq(0.1, 3, length.out = 300),
    dens  = dlnorm(seq(0.1, 3, length.out = 300), meanlog = 0, sdlog = 0.3)
  )
) |>
  mutate(param = factor(param,
                        levels = c("beta[0]", "beta[1]", "sigma^2")))

p_dens <- mcmc_df |>
  filter(phase == "Post-burn-in") |>
  ggplot(aes(value)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 40, fill = "#3a7abf", colour = "white", alpha = 0.5) +
  geom_density(colour = "#3a7abf", linewidth = 0.9) +
  geom_line(data = prior_curves,
            aes(x = x, y = dens),
            colour = "#e05c5c", linewidth = 1, linetype = "dashed",
            inherit.aes = FALSE) +
  geom_vline(data = true_vals,
             aes(xintercept = truth),
             colour = "black", linetype = "dotted", linewidth = 0.9) +
  facet_wrap(~ param, scales = "free", labeller = label_parsed) +
  labs(
    title    = "Prior (red dashed) vs Posterior (blue)  |  True value (black dotted)",
    subtitle = "Posterior should be narrower than prior and centered near the truth",
    x = "Parameter value", y = "Density"
  ) +
  lab_theme()

(p_trace / p_dens) +
  plot_layout(heights = c(3, 2)) +
  plot_annotation(
    title = "MCMC Diagnostics",
    theme = theme(plot.title = element_text(
                    face = "bold", hjust = 0.5, size = 16))
  )
```

------------------------------------------------------------------------

# Step 5: Posterior Summaries

The table below summarizes the **marginal posterior distribution** for
each parameter using the post-burn-in draws.

**Column guide:**

| Column | Meaning |
|------------------------------------|------------------------------------|
| Mean | Posterior mean — our best single-number estimate |
| SD | Posterior standard deviation — analogous to a standard error |
| CrI Lower / Upper | 95% equal-tail credible interval — the central 95% of posterior mass |

> **Credible interval vs. confidence interval:** A **95% Bayesian
> credible interval** $[a, b]$ means: *Given the data, there is a 95%
> probability that the true parameter lies in* $[a, b]$. A **95%
> frequentist confidence interval** does NOT have this interpretation.
> It means: if we repeated the experiment many times, 95% of the
> resulting intervals would contain the true value.

```{r step5-summaries}
summary_tbl <- bind_rows(
  posterior_summary(beta0_chain,  burn) |> mutate(Parameter = "\u03b2\u2080"),
  posterior_summary(beta1_chain,  burn) |> mutate(Parameter = "\u03b2\u2081"),
  posterior_summary(sigma2_chain, burn) |> mutate(Parameter = "\u03c3\u00b2")
) |>
  select(Parameter, everything())

knitr::kable(
  summary_tbl, align = "lrrrrr",
  caption = "Posterior summaries \u2014 post-burn-in samples only"
)
```

------------------------------------------------------------------------

# Step 6: Posterior Predictive Distribution

The **posterior predictive distribution** answers the question: *For a
new day with* $x^* = 5$ customers, what revenue should we expect?

Unlike OLS, which gives a single point prediction plus a standard error,
the Bayesian approach gives a **full predictive distribution** that
propagates *two* sources of uncertainty:

1.  **Parameter uncertainty** — we don't know the exact values of
    $\beta_0$, $\beta_1$, $\sigma^2$. We represent this by drawing
    $(\beta_0^{(s)}, \beta_1^{(s)}, \sigma^{2(s)})$ from the posterior.
2.  **Observation noise** — even if we knew the true parameters, future
    revenues would still vary randomly. We capture this by adding
    $\varepsilon^{(s)} \sim N(0, \sigma^{2(s)})$ to each prediction.

The predictive draw is therefore:
$$\tilde{y}^{(s)} = \beta_0^{(s)} + \beta_1^{(s)} x^* + \varepsilon^{(s)},
\quad \varepsilon^{(s)} \sim N(0,\, \sigma^{2(s)})$$

This is implemented by:

1.  Sampling `n_samp = 1000` random indices from the post-burn-in chain
    (with replacement, to handle any chain length).
2.  Using those indexed draws of $(\beta_0, \beta_1, \sigma^2)$ as if
    they were independent posterior samples.
3.  Adding independent normal noise scaled by $\sqrt{\sigma^{2(s)}}$.

```{r step6-predictive, fig.height=9}
n_samp  <- 1000   # number of predictive samples
x_star  <- 5      # new predictor value: 5 customers
x_curve <- seq(0, 10, length.out = 100)   # grid for the ribbon plot

# Sample indices from post-burn-in chain (with replacement)
post_idx <- sample(post_burn:n_iter, n_samp, replace = TRUE)

# ── Point prediction at x* = 5 ────────────────────────────────────────
# For each posterior draw, compute:
#   mean component  +  random noise
y_star <- beta0_chain[post_idx] +
          beta1_chain[post_idx] * x_star +
          rnorm(n_samp, 0, sqrt(sigma2_chain[post_idx]))

# 95% equal-tail credible interval for the prediction
cri_95 <- round(quantile(y_star, c(0.025, 0.975)), 2)

# ── Histogram of predictive distribution ──────────────────────────────
p_hist <- ggplot(tibble(y_pred = y_star), aes(y_pred)) +
  geom_histogram(aes(y = after_stat(density)),
                 bins = 35, fill = "#667eea",
                 colour = "white", alpha = 0.85) +
  geom_density(colour = "#3a3a8c", linewidth = 1) +
  geom_vline(xintercept = mean(y_star),
             colour = "#d62728", linetype = "dashed", linewidth = 1) +
  annotate("text",
           x = mean(y_star) + 0.15, y = Inf,
           label = paste0("mean = ", round(mean(y_star), 1)),
           vjust = 2, hjust = 0, colour = "#d62728", size = 3.8) +
  labs(
    title    = paste0("Predictive distribution  |  x \u2217 = ", x_star),
    subtitle = paste0("Mean: ", round(mean(y_star), 2),
                      "   95% CrI: [", cri_95[1], ", ", cri_95[2], "]"),
    x = "Predicted revenue ($)", y = "Density"
  ) +
  lab_theme()

# ── Full predictive ribbon across x = 0 to 10 ─────────────────────────
# For each x value on the grid, draw 1000 predictive samples and compute
# the mean and 2.5th / 97.5th percentiles. This gives the "credible band."
curve_df <- map_dfr(seq_along(x_curve), function(j) {
  sims <- beta0_chain[post_idx] +
          beta1_chain[post_idx] * x_curve[j] +
          rnorm(n_samp, 0, sqrt(sigma2_chain[post_idx]))
  tibble(
    x    = x_curve[j],
    mean = mean(sims),
    lo   = quantile(sims, 0.025),
    hi   = quantile(sims, 0.975)
  )
})

p_ribbon <- ggplot() +
  geom_ribbon(data = curve_df,
              aes(x = x, ymin = lo, ymax = hi),
              fill = "#4ECDC4", alpha = 0.35) +
  geom_line(data = curve_df,
            aes(x = x, y = mean),
            colour = "#2E86AB", linewidth = 1.4) +
  geom_point(data = coffee,
             aes(x = customers, y = revenue),
             colour = "#FF6B6B", size = 2.5, alpha = 0.80) +
  geom_smooth(data = coffee,
              aes(x = customers, y = revenue),
              method = "lm", se = FALSE,
              colour = "grey50", linewidth = 0.8, linetype = "dotted") +
  # Highlight the x* = 5 prediction
  geom_point(aes(x = x_star, y = mean(y_star)),
             shape = 23, fill = "#f0e442",
             colour = "black", size = 4) +
  annotate("text",
           x = x_star + 0.2, y = mean(y_star) + 1.2,
           label = paste0("x \u2217 = ", x_star,
                          "\n\u0233 = ", round(mean(y_star), 1)),
           size = 3.6, hjust = 0) +
  labs(
    title    = "Posterior Predictive \u2014 Full Curve",
    subtitle = "Teal ribbon: 95% credible band  |  Red: data  |  Grey dotted: OLS",
    x = "Customers", y = "Revenue ($)"
  ) +
  lab_theme()

p_hist / p_ribbon
```

------------------------------------------------------------------------
