---
title: "Lab 05: The Coffee Shop Oracle — Lab Activity"
subtitle: "MA214 Applied Statistics — Bayesian Linear Regression"
author: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.width  = 10,
  fig.height = 6,
  fig.align  = "center",
  out.width  = "95%"
)
library(tidyverse)
library(patchwork)
```

------------------------------------------------------------------------

## Background

**BU Coffee Shop** wants to predict **daily revenue** from number of
customers. You have been given `Coffee_df.csv`, a dataset recording 60
days of observations. The Bayesian linear regression model is:

$$y_i = \beta_0 + \beta_1\,x_i + \varepsilon_i,
\qquad \varepsilon_i \overset{iid}{\sim} N(0,\,\sigma^2)$$

| Symbol     | Meaning                                    |
|------------|--------------------------------------------|
| $y_i$      | Daily revenue (\$)                         |
| $x_i$      | Customers per day                          |
| $\beta_0$  | Baseline revenue when $x = 0$              |
| $\beta_1$  | Extra revenue per additional customer      |
| $\sigma^2$ | Day-to-day unexplained variance in revenue |

------------------------------------------------------------------------

## The Model

**Likelihood** — how revenue arises given the parameters:
$$y_i \mid \beta_0, \beta_1, \sigma^2
  \;\sim\; N\!\left(\beta_0 + \beta_1\,x_i,\; \sigma^2\right)$$

**Priors** — our domain knowledge before seeing the data:
$$\beta_0 \sim N(5,\,2^2), \qquad
\beta_1 \sim N(2,\,0.5^2), \qquad
\sigma^2 \sim \text{Log-Normal}(\log 1,\,0.3^2)$$

> **Prior intuition:**
>
> - $\beta_0 \sim N(5, 4)$: baseline revenue around \$5 with SD \$2.
> - $\beta_1 \sim N(2, 0.25)$: each extra customer brings roughly \$2,
>   tightly centered (95% prior range ≈ [\$1.02, \$2.98]).
> - $\sigma^2 \sim \text{LogNormal}(0, 0.09)$: prior median = 1, 95%
>   range ≈ [0.55, 1.80] — noise is modest.

**Log-posterior** (by Bayes' theorem, dropping the normalizing
constant): $$\log p(\beta_0, \beta_1, \sigma^2 \mid \mathbf{y})
  = \underbrace{\sum_{i=1}^{n} \log N(y_i;\,\mu_i,\,\sigma^2)}_{\text{log-likelihood}}
  + \underbrace{\log p(\beta_0) + \log p(\beta_1) + \log p(\sigma^2)}_{\text{log-priors}}$$

------------------------------------------------------------------------

## Activity

You will complete **Tasks 1, 2, and 4** in `lab5.R` and submit it to
GradeScope. Task 3 (MCMC) is provided for you — read it carefully to
understand what the chains contain before moving to Task 4.

**You are not supposed to use GenAI for this activity !!!**

------------------------------------------------------------------------

### Task 1 — Load the data

Load `Coffee_df.csv` into a data frame called `df`.

```{r task1-load}
library(dplyr)
library(ggplot2)
library(readr)

# ── Task 1 ────────────────────────────────────────────────────────────
df <- read.csv("Coffee_df.csv")

glimpse(df)
```

**Quick plot** — does revenue increase with customers as expected?

```{r task1-plot}
ggplot(df, aes(x = customers, y = revenue)) +
  geom_point(colour = "#1f77b4", size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = TRUE,
              colour = "#d62728", fill = "#d62728", alpha = 0.15) +
  labs(
    title    = "Daily Revenue vs. Customers",
    subtitle = paste0("n = ", nrow(df), " days  |  OLS line shown in red"),
    x = "Customers per day",
    y = "Revenue ($)"
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title    = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, colour = "grey40"))
```

> **Question 1.** From the plot, roughly how much does one extra
> customer add to daily revenue? Does the relationship look linear?

*Your answer here:*

------------------------------------------------------------------------

### Task 2 — Define `log_posterior()`

Write `log_posterior(beta0, beta1, sigma2)` returning
$\log p(\beta_0, \beta_1, \sigma^2 \mid \mathbf{y})$.

**Checklist:**

- [ ] Returns `-Inf` when `sigma2 <= 0`
- [ ] Log-likelihood:
  `sum(dnorm(y, mean = mu, sd = sqrt(sigma2), log = TRUE))`
- [ ] Priors: `dnorm(beta0, 5, 2, log = TRUE)` ·
  `dnorm(beta1, 2, 0.5, log = TRUE)` ·
  `dlnorm(sigma2, meanlog = 0, sdlog = 0.3, log = TRUE)`
- [ ] Returns a single finite number at valid inputs

```{r task2-logpost}
x <- df$customers
y <- df$revenue

# ── Task 2 ────────────────────────────────────────────────────────────
log_posterior <- function(beta0, beta1, sigma2) {

  # (a) Guard: return -Inf if sigma2 is non-positive
  if (sigma2 <= 0) return(-Inf)

  # (b) Log-likelihood: sum of log N(y_i; mu_i, sigma2)
  mu      <- ________________________
  log_lik <- ________________________

  # (c) Log-priors
  log_p_beta0  <- ________________________   # N(5, 2^2)
  log_p_beta1  <- ________________________   # N(2, 0.5^2)
  log_p_sigma2 <- ________________________   # LogNormal(log1, 0.3^2)

  # (d) Return log-posterior = log-likelihood + log-priors
  ________________________
}
```

**Self-check** — all four lines must print `TRUE` before continuing.

```{r task2-check}
cat("Returns -Inf for sigma2 = 0:", log_posterior(5, 2,  0) == -Inf, "\n")
cat("Returns -Inf for sigma2 < 0:", log_posterior(5, 2, -1) == -Inf, "\n")
cat("Finite at (5, 2, 1):        ", is.finite(log_posterior(5, 2, 1)), "\n")
cat("True > bad params:          ",
    log_posterior(8.5, 1.75, 38.44) > log_posterior(100, -5, 0.001), "\n")
```

> **Question 2.** Why must `log_posterior()` return `-Inf` (not `NaN` or
> an error) when `sigma2 <= 0`? What happens inside the MH loop if it
> returns `-Inf`?

*Your answer here:*

------------------------------------------------------------------------

### Task 3 — MCMC (provided — do not modify)

The Metropolis-Hastings sampler below draws from the posterior using
**component-wise updates** (one parameter at a time). Run it, inspect
the acceptance rates and trace plots, then proceed to Task 4.

```{r task3-mcmc}
# ── MCMC settings ─────────────────────────────────────────────────────
set.seed(214214)
n_iter <- 6000
burn   <- 1000

beta0_chain  <- numeric(n_iter)
beta1_chain  <- numeric(n_iter)
sigma2_chain <- numeric(n_iter)

beta0_chain[1]  <- 5   # starting at prior mean
beta1_chain[1]  <- 2   # starting at prior mean
sigma2_chain[1] <- 1   # starting at prior median

# Proposal SDs — pre-tuned for ~40% per-parameter acceptance
prop_sd <- c(beta0 = 1.2, beta1 = 0.02, log_sigma2 = 0.15)
acc     <- c(b0 = 0L, b1 = 0L, s2 = 0L)

for (i in seq(2, n_iter)) {
  b0_cur <- beta0_chain[i - 1]
  b1_cur <- beta1_chain[i - 1]
  s2_cur <- sigma2_chain[i - 1]

  # Update beta0
  b0_prop <- rnorm(1, b0_cur, prop_sd["beta0"])
  log_a   <- log_posterior(b0_prop, b1_cur, s2_cur) -
             log_posterior(b0_cur,  b1_cur, s2_cur)
  if (log(runif(1)) < log_a) {
    beta0_chain[i] <- b0_prop; acc["b0"] <- acc["b0"] + 1L
  } else { beta0_chain[i] <- b0_cur }

  # Update beta1 (using updated beta0)
  b0_cur  <- beta0_chain[i]
  b1_prop <- rnorm(1, b1_cur, prop_sd["beta1"])
  log_a   <- log_posterior(b0_cur, b1_prop, s2_cur) -
             log_posterior(b0_cur, b1_cur,  s2_cur)
  if (log(runif(1)) < log_a) {
    beta1_chain[i] <- b1_prop; acc["b1"] <- acc["b1"] + 1L
  } else { beta1_chain[i] <- b1_cur }

  # Update sigma2 on log scale (using updated beta0, beta1)
  b1_cur  <- beta1_chain[i]
  s2_prop <- exp(rnorm(1, log(s2_cur), prop_sd["log_sigma2"]))
  log_a   <- log_posterior(b0_cur, b1_cur, s2_prop) -
             log_posterior(b0_cur, b1_cur, s2_cur)
  if (log(runif(1)) < log_a) {
    sigma2_chain[i] <- s2_prop; acc["s2"] <- acc["s2"] + 1L
  } else { sigma2_chain[i] <- s2_cur }
}

cat("Acceptance rates — beta0:", round(acc["b0"] / (n_iter-1) * 100, 1), "%",
    "| beta1:", round(acc["b1"] / (n_iter-1) * 100, 1), "%",
    "| sigma2:", round(acc["s2"] / (n_iter-1) * 100, 1), "%\n")
```

**Trace plots** — verify convergence before using the chains.

```{r task3-trace, fig.height=8}
post_burn <- burn + 1

tibble(iter = 1:n_iter,
       beta0 = beta0_chain, beta1 = beta1_chain, sigma2 = sigma2_chain) |>
  pivot_longer(-iter, names_to = "param", values_to = "value") |>
  mutate(
    phase = if_else(iter <= burn, "Burn-in", "Post-burn-in"),
    param = factor(param,
                   levels = c("beta0", "beta1", "sigma2"),
                   labels = c("beta[0]", "beta[1]", "sigma^2"))
  ) |>
  ggplot(aes(iter, value, colour = phase)) +
  geom_line(linewidth = 0.35, alpha = 0.9) +
  facet_wrap(~ param, scales = "free_y", ncol = 1, labeller = label_parsed) +
  scale_colour_manual(
    values = c("Burn-in" = "#e05c5c", "Post-burn-in" = "#3a7abf")
  ) +
  labs(title    = "Trace Plots",
       subtitle = "Red = burn-in (discarded)  |  Blue = post-burn-in (kept)",
       x = "Iteration", y = "Value", colour = NULL) +
  theme_minimal(base_size = 13) +
  theme(plot.title    = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, colour = "grey40"),
        legend.position = "bottom")
```

Post-burn-in draws — these are the samples you will use in Task 4.

```{r task3-post}
post_b0 <- beta0_chain[post_burn:n_iter]   # 5000 posterior draws for beta0
post_b1 <- beta1_chain[post_burn:n_iter]   # 5000 posterior draws for beta1
post_s2 <- sigma2_chain[post_burn:n_iter]  # 5000 posterior draws for sigma2

cat("Posterior means:\n")
cat("  beta0: ",  round(mean(post_b0), 3), "\n")
cat("  beta1: ",  round(mean(post_b1), 3), "\n")
cat("  sigma2:",  round(mean(post_s2), 3), "\n")

# beta1_CI — autograder target from this lab
beta1_CI <- quantile(post_b1, c(0.025, 0.975))
cat("\n95% Credible Interval for beta1: [",
    round(beta1_CI[1], 3), ",", round(beta1_CI[2], 3), "]\n")
```

------------------------------------------------------------------------

### Task 4 — Posterior Predictive Mean

Now that you have posterior draws, predict revenue for a new day with
$x^* = 3$ customers.

**The posterior predictive distribution** integrates out all parameter
uncertainty:

$$p(\tilde{y} \mid \mathbf{y},\, x^*) =
  \int p(\tilde{y} \mid \beta_0, \beta_1, \sigma^2,\, x^*)\;
  p(\beta_0, \beta_1, \sigma^2 \mid \mathbf{y})\;
  d\beta_0\, d\beta_1\, d\sigma^2$$

Approximate this with Monte Carlo: for each of `n_samp = 1000` randomly
sampled posterior draws $(\beta_0^{(s)}, \beta_1^{(s)},
\sigma^{2(s)})$, generate one predictive observation:

$$\tilde{y}^{(s)} = \underbrace{\beta_0^{(s)} + \beta_1^{(s)}\,x^*}_{\text{mean component}}
  + \underbrace{\varepsilon^{(s)}}_{\text{noise}},
  \qquad \varepsilon^{(s)} \sim N(0,\, \sigma^{2(s)})$$

The **posterior predictive mean** is $\texttt{post\_mean} =
\frac{1}{1000}\sum_{s=1}^{1000} \tilde{y}^{(s)}$.

> **Two sources of uncertainty:** (1) **Parameter uncertainty** — we
> don't know the exact $(\beta_0, \beta_1, \sigma^2)$, so we draw from
> the posterior. (2) **Observation noise** — even with known parameters,
> tomorrow's revenue varies randomly; we add $\varepsilon^{(s)}$ for
> this.

```{r task4-predictive}
# ── Task 4 ────────────────────────────────────────────────────────────
x_star <- 3      # new day: 3 customers
n_samp <- 1000   # number of Monte Carlo draws

# Step 1: randomly sample n_samp indices from the 5000 posterior draws
set.seed(42)
idx <- sample(length(post_b1), n_samp, replace = TRUE)

# Step 2: compute one predictive draw per sampled parameter set
#   y_pred[s] = beta0[s] + beta1[s] * x_star + rnorm(1, 0, sqrt(sigma2[s]))
y_pred <- ________________________   # hint: vectorise over idx

# Step 3: posterior predictive mean
post_mean <- ________________________

cat("x* =", x_star, "customers\n")
cat("post_mean = $", round(post_mean, 3), "\n")
```

**Visualise the predictive distribution:**

```{r task4-plot}
ggplot(data.frame(y_pred = y_pred), aes(x = y_pred)) +
  geom_histogram(aes(y = after_stat(density)), bins = 35,
                 fill = "#667eea", colour = "white", alpha = 0.85) +
  geom_density(colour = "#3a3a8c", linewidth = 1) +
  geom_vline(xintercept = post_mean,
             colour = "#d62728", linetype = "dashed", linewidth = 1) +
  annotate("text",
           x = post_mean + 0.3, y = Inf,
           label = paste0("post_mean = $", round(post_mean, 2)),
           vjust = 2, hjust = 0, colour = "#d62728", size = 4) +
  labs(
    title    = paste0("Posterior Predictive Distribution  |  x* = ", x_star),
    subtitle = "Each draw propagates parameter uncertainty and observation noise",
    x = "Predicted Revenue ($)",
    y = "Density"
  ) +
  theme_minimal(base_size = 13) +
  theme(plot.title    = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, colour = "grey40"))
```

> **Question 4a.** Write a one-sentence interpretation of `post_mean` in
> the context of the coffee shop (use dollar amounts).

*Your answer here:*

> **Question 4b.** The true data-generating mean at $x^* = 3$ is
> $8.5 + 1.75 \times 3 = \$13.75$. Why might your `post_mean` differ
> from this value?

*Your answer here:*

> **Question 4c.** If you removed the `rnorm(...)` noise term from Step
> 2 and just used $\tilde{y}^{(s)} = \beta_0^{(s)} +
> \beta_1^{(s)}\,x^*$, would `post_mean` change much? What would the
> predictive distribution look like — narrower or wider?

*Your answer here:*

------------------------------------------------------------------------

## Posterior Summary

```{r summary-table}
summary_df <- data.frame(
  Parameter  = c("beta0", "beta1", "sigma2"),
  `Post Mean` = round(c(mean(post_b0), mean(post_b1), mean(post_s2)), 3),
  `Post SD`   = round(c(sd(post_b0),   sd(post_b1),   sd(post_s2)),   3),
  `CrI 2.5%`  = round(c(quantile(post_b0, 0.025),
                         beta1_CI["2.5%"],
                         quantile(post_s2, 0.025)), 3),
  `CrI 97.5%` = round(c(quantile(post_b0, 0.975),
                         beta1_CI["97.5%"],
                         quantile(post_s2, 0.975)), 3),
  check.names = FALSE
)
knitr::kable(summary_df, caption = "Posterior summaries — post-burn-in samples")
```

```{r summary-densities, fig.height=4}
tibble(beta0 = post_b0, beta1 = post_b1, sigma2 = post_s2) |>
  pivot_longer(everything(), names_to = "param", values_to = "value") |>
  mutate(param = factor(param,
                        levels = c("beta0", "beta1", "sigma2"),
                        labels = c("beta[0]", "beta[1]", "sigma^2"))) |>
  ggplot(aes(value, fill = param)) +
  geom_histogram(aes(y = after_stat(density)), bins = 40,
                 colour = "white", alpha = 0.7) +
  geom_density(colour = "black", linewidth = 0.8) +
  facet_wrap(~ param, scales = "free", labeller = label_parsed) +
  scale_fill_brewer(palette = "Set2") +
  labs(title    = "Posterior Marginal Distributions",
       subtitle = "Post-burn-in samples only",
       x = "Parameter value", y = "Density") +
  theme_minimal(base_size = 13) +
  theme(plot.title    = element_text(face = "bold", hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, colour = "grey40"),
        legend.position = "none")
```

------------------------------------------------------------------------

## GradeScope Submission

> **AutoGrader requirements:**
>
> - `df` — dataframe using `read.csv()`
> - `log_posterior()` — function
> - `beta1_CI`— a vector of size 2
> - `post_mean` — numeric value

```{r postlab-activity}
# ============================================================
# SUBMISSION INSTRUCTIONS — READ CAREFULLY
# ============================================================
# Submit lab5.R to GradeScope, NOT this Rmd file.
#
# Your lab5.R must define exactly these four objects:
#
#   df             — data frame from read.csv("Coffee_df.csv")
#   log_posterior  — function(beta0, beta1, sigma2)
#   beta1_CI       — named numeric vector, length 2:
#                    quantile(post_b1, c(0.025, 0.975))
#   post_mean      — numeric scalar: mean of 1000 predictive
#                    draws at x_star = 3, using set.seed(42)
#
# Rules:
#   - Do NOT use setwd().
#   - Only import: dplyr, ggplot2, readr.
#   - MCMC: set.seed(214214), n_iter=6000, burn=1000,
#     starting values beta0=5, beta1=2, sigma2=1.
#   - Task 4: x_star=3, n_samp=1000, set.seed(42).
#   - Run lab5.R completely before submitting.
#   - Pass = 4/4 on the autograder.
# ============================================================

library(dplyr)
library(ggplot2)
library(readr)

df <- read.csv("Coffee_df.csv") # [Task 1] df

# [Task 2] copy your log_posterior() here

# [Task 3] copy the MCMC block here (provided above — do not modify)

# [Task 4] copy your y_pred and post_mean code here
```
