---
title: "Lab 03: Bayes' Rule"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

------------------------------------------------------------------------

# Coin Flip Game

There are five different coin types. Each coin has different probability
of getting head. See the table below.

| Coin Type                   | A    | B    | C    | D    | E    |
|-----------------------------|------|------|------|------|------|
| Probability of getting head | 0.15 | 0.30 | 0.50 | 0.70 | 0.90 |

With each different type of coin, all the flipped results were written
in the csv file. Imagine a dealer gives you a different type of coin.

Your job is to figure out which type the coin is at the end by using
Bayesian knowledge.

------------------------------------------------------------------------

## Step 0.1: Choose the Study

Run the code below. It randomly assigns you a coin type to study. Load
the corresponding CSV file in Step 2.1.

```{r}
data_types <- seq(1:10)
your_study <- sample(data_types, 1)
cat("Your data is:", your_study)
```

> **Note:** Your assigned number corresponds to a CSV file (e.g., number
> `3` → `coin_03.csv`). Files run from `coin_01.csv` to `coin_10.csv`.

------------------------------------------------------------------------

## Step 0.2: Theory Review

Bayes' Rule lets us update our beliefs about an unknown parameter after
observing data:

$$P(A \mid B) = \frac{P(A \cap B)}{P(B)} = \frac{P(B \mid A)\, P(A)}{P(B)}$$

In Bayesian language:

$$\text{Posterior} = \frac{\text{Likelihood} \times \text{Prior}}{\text{Normalizing Constant}}$$

------------------------------------------------------------------------

## Step 0.3: Load Data

Replace the filename matching your assigned coin type.

```{r load-data}
my_coin_file <- "coin_01.csv"   #TODO: change to your assigned file
coin_df      <- read.csv(my_coin_file)
head(coin_df, 10)
```

```{r data-summary}
n_flips     <- nrow(coin_df)
n_heads     <- sum(coin_df$result)
n_tails     <- n_flips - n_heads
sample_prop <- round(n_heads / n_flips, 3)

cat("File:             ", my_coin_file, "\n")
cat("Total flips:      ", n_flips,      "\n")
cat("Heads:            ", n_heads,      "\n")
cat("Tails:            ", n_tails,      "\n")
cat("Sample proportion:", sample_prop,  "\n")
```

------------------------------------------------------------------------

## Step 0.4: Setting Up the Notation and Prior

#### Notation

In this lab, we are trying to identify **which coin type** the dealer
gave us. The five coin types have different probabilities of landing
heads. We label these known probabilities as:

$$p_j \quad \text{for } j = A, B, C, D, E$$

Specifically:

| Coin type $j$ | $p_j = P(\text{Heads} \mid \text{type } j)$ |
|:-------------:|:-------------------------------------------:|
|       A       |                    0.15                     |
|       B       |                    0.30                     |
|       C       |                    0.50                     |
|       D       |                    0.70                     |
|       E       |                    0.90                     |

> **Key distinction:** $p_j$ is the *known* probability of heads
> **given** coin type $j$. It is a fixed property of each coin — not
> something we are estimating. What we are estimating is **which type**
> $j$ the dealer gave us — that is the unknown.

The unknown we are inferring is the **coin type**
$j \in \{A, B, C, D, E\}$. We assign probabilities to each type using
**Bayes' Rule**:

$$P(\pi_j \mid \text{data}) = \frac{P(\text{data} \mid \pi_j) \cdot P(\pi_j)}{C}$$

where:

-   $P(\pi_j)$ is the **prior**: our belief about having coin type $j$
    *before* seeing any flips

-   $P(\text{data} \mid \pi_j)$ is the **likelihood**: how probable the
    observed flips are *if* the coin were type $j$

-   $C = \sum_{j} P(\text{data} \mid \pi_j)\, P(\pi_j)$ is the
    **normalizing constant**: ensures the posterior sums to 1

-   $P(\pi_j \mid \text{data})$ is the **posterior**: our updated belief
    about the coin type *after* seeing the flips

## Step 1. First approach : Using the full data

#### Exercise 1. Initial Prior

```{r prior}
p_js         <- c(0.15, 0.30, 0.50, 0.70, 0.90)   # known head prob for each type
coin_labels <- c("A",  "B",  "C",  "D",  "E")
prior_probs <- # TODO : Set up the initial priors (size 5 vector)

prior_df <- tibble(
  type  = coin_labels,
  p_j   = p_js,
  prior = prior_probs
)

prior_df
```

```{r prior-plot, fig.height=3.5, fig.width=6}
# visulization
prior_df |>
  ggplot(aes(x = paste0(type, " (p_j=", p_j, ")"), y = prior)) +
  geom_col(fill = "steelblue", alpha = 0.85, width = 0.55) +
  geom_text(aes(label = prior), vjust = -0.5, size = 4) +
  labs(title = "Prior Distribution over Coin Type",
       subtitle = "Initial Priors ",
       x = "Coin Type", y = expression(P(j))) +
  ylim(0, 0.35) +
  theme_minimal(base_size = 13)
```

------------------------------------------------------------------------

### Step 1.2: Compute the Likelihood

$$P(j \mid \text{data}) = \frac{P(\text{data} \mid j) \cdot P(j)}{C}$$

where $C = \sum_j P(data∣j)P(j)$ is the normalizing constant.

For $n$ flips with $x$ heads, the **Binomial likelihood** for coin type
$j$ — i.e., the probability of observing $x$ heads if the coin were type
$j$ with known head probability $p_j$ — is:

$$P(X = x \mid \pi_j) = \binom{n}{x}\, p_j^{\,x}\, (1 - p_j)^{n - x}$$

Note that we plug in $p_j$ (the *known* head probability for type $j$) —
not an unknown $\pi$ we are estimating.

```         
dbinom(x, size, prob)
```

```         
Description
Density, distribution function, quantile function and random generation for the binomial distribution with parameters size and prob.

This is conventionally interpreted as the number of ‘successes’ in size trials.
```

#### Exercise 2. Calculate the likelihood of having the number of heads from the data and number of flips given the each type of coins.

```{r likelihood}
# Compute likelihood P(data | j) for each coin type j
# using p_j (the known head probability for that type)
prior_df <- prior_df |>
  mutate(
    likelihood = #TODO : likelihood of those 5 types of coins (size 5 vector)
  )

prior_df
```

#### Exercise 3. Calculate the normalizing constant and the posterior probability for each p_j.

```{r posterior-batch}
# Apply Bayes' Rule: P(j | data) = P(data | j) * P(j) / C
update_df <- prior_df |>
  mutate(
    numerator = #TODO: P(data | j) * P(j) (size 5 vector)
    posterior = #TODO: Scalar # Normalizing constant C
  )

cat("Normalizing constant C =", round(sum(update_df$numerator), 7), "\n\n")
update_df
```

```{r posterior-plot, fig.height=4, fig.width=7}
# visulization
update_df |>
  select(type, p_j, prior, posterior) |>
  pivot_longer(c(prior, posterior), names_to = "dist", values_to = "prob") |>
  mutate(
    dist  = factor(dist, levels = c("prior", "posterior"),
                   labels = c("Prior", "Posterior (all flips)")),
    label = paste0(type, " (p_j=", p_j, ")")
  ) |>
  ggplot(aes(x = label, y = prob, fill = dist)) +
  geom_col(position = "dodge", alpha = 0.85, width = 0.6) +
  scale_fill_manual(values = c("Prior" = "steelblue",
                               "Posterior (all flips)" = "tomato")) +
  labs(title = paste0("Prior vs. Posterior — ", my_coin_file),
       subtitle = paste0(n_heads, " heads out of ", n_flips, " flips"),
       x = "Coin Type", y = "Probability", fill = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```

#### Exercise 4. Based on the batch posterior above, which coin type do you think the dealer gave you? State the type, its posterior probability, and explain why in one sentence.

**Answer:**

------------------------------------------------------------------------

## Step 2: Sequential Updating

Instead of applying Bayes' Rule to all flips at once, we can update the
posterior **one flip at a time**. Each flip's posterior becomes the
prior for the next flip.

For one flips, the **Binomial likelihood** for coin type $j$ becomes
**Bernoulli likelihood**— i.e., the probability of observing a head if
the coin were type $j$ with known head probability $p_j$ — is:

$$P(X = x \mid \pi_j) =  p_j^{\,x}\, (1 - p_j)^{1 - x}$$

where $x =1$ if it is head, $0$ otherwise.

Note that we plug in $p_j$ (the *known* head probability for type $j$) —
not an unknown $\pi$ we are estimating.

```         
dbinom(x, 1, prob) # You can use the same function in R
```

------------------------------------------------------------------------

### Step 2.1: Fill Out the Table (First 3 Flips)

### Exercise 5. The table below shows the prior × likelihood → posterior calculation for

the first **3 flips** manually, so you can see the mechanics before
automating it.

```{r manual-3-flips 0}
flips <- coin_df$result   # 1 = Heads, 0 = Tails

cat("First 3 flips:", flips[1], flips[2], flips[3],
    "(1=H, 0=T)\n\n")

# --- Flip 0: Prior ---
p0 <- #TODO: starting distribution before any flips

```

```{r manual-3-flips 1}
# --- Flip 1 ---
lk1 <- #TODO: P(flip1 | each coin type)
p1  <- #TODO: posterior after flip 1
```

```{r manual-3-flips 2}
# --- Flip 2 ---
lk2 <- #TODO: P(flip2 | each coin type)
p2  <- #TODO: posterior after flip 2

```

```{r manual-3-flips 3}
# --- Flip 2 ---
lk3 <- #TODO: P(flip3 | each coin type)
p3  <- #TODO: posterior after flip 3

```

```{r manual-3-flips 4}

# Display as a table
manual_table <- tibble(
  Stage      = c("Prior (flip 0)", "After flip 1", "After flip 2", "After flip 3"),
  Flip_result = c(NA, flips[1], flips[2], flips[3]),
  `A (0.15)` = round(c(p0[1], p1[1], p2[1], p3[1]), 4),
  `B (0.30)` = round(c(p0[2], p1[2], p2[2], p3[2]), 4),
  `C (0.50)` = round(c(p0[3], p1[3], p2[3], p3[3]), 4),
  `D (0.70)` = round(c(p0[4], p1[4], p2[4], p3[4]), 4),
  `E (0.90)` = round(c(p0[5], p1[5], p2[5], p3[5]), 4)
)

manual_table

```

------------------------------------------------------------------------

### Step 2.2: Implement Full Sequential Updating

### Exercise 6. Now automate the update across all flips:

```{r sequential-update}
# Matrix to store posterior at each step: rows = flips+1, cols = 5 coin types
post_mat      <- matrix(NA, nrow = n_flips + 1, ncol = 5)
post_mat[1, ] <- prior_probs   # row 1 = prior (before any flips)

for (i in 1:n_flips) {
  lk              <-  #TODO: P(flip_i | each type)
  unnorm          <- #TODO: prior * likelihood (unnormalized)
  C <-              #TODO: normalizing constant
  post_mat[i+1, ] <- #TODO:
}

# Posterior mean: weighted average using the known p_j values
post_mean_seq <- apply(post_mat, 1, function(post) sum(post * p_js))

cat("Sequential updating complete.\n")
cat("Posterior after ALL", n_flips, "flips:\n")
tibble(type = coin_labels, p_j = p_js,
       posterior = round(post_mat[n_flips + 1, ], 4))
```

------------------------------------------------------------------------

### Step 2.3: Snapshot Table and Trajectory

```{r snapshot-table}
snapshots <- c(0, 5, 10, 20, n_flips)

map_dfr(snapshots, function(k) {
  tibble(
    `After flips` = k,
    `A (0.15)`    = round(post_mat[k+1, 1], 4),
    `B (0.30)`    = round(post_mat[k+1, 2], 4),
    `C (0.50)`    = round(post_mat[k+1, 3], 4),
    `D (0.70)`    = round(post_mat[k+1, 4], 4),
    `E (0.90)`    = round(post_mat[k+1, 5], 4)
  )
})
```

```{r trajectory-plot, fig.height=4, fig.width=8}
tibble(flip = 0:n_flips, post_mean = post_mean_seq) |>
  ggplot(aes(x = flip, y = post_mean)) +
  geom_line(color = "steelblue", linewidth = 1.1) +
  geom_point(data = tibble(flip = snapshots,
                           post_mean = post_mean_seq[snapshots + 1]),
             color = "tomato", size = 3) +
  geom_hline(yintercept = 0.50, linetype = "dashed",
             color = "gray50", linewidth = 0.8) +
  annotate("text", x = n_flips * 0.82, y = 0.53,
           label = "Fair coin C (p_j = 0.50)", color = "gray50", size = 3.5) +
  labs(title = "Posterior Mean After Each Flip",
       subtitle = "Red points = snapshot stages | Does it drift above or below 0.50?",
       x = "Number of flips",
       y = expression("Posterior mean of " ~ p[j])) +
  ylim(0.10, 0.95) +
  theme_minimal(base_size = 13)
```

------------------------------------------------------------------------

### Exercise 7. Are the Posteriors the Same?

Compare the batch posterior (all flips at once, from Step 1) to the
sequential posterior (flip by flip, from Step 2). Do they match? then
what does it tell you?

**Answer:**

------------------------------------------------------------------------

### Exercise 8. What Is Your Final Guess?

**Answer:**

------------------------------------------------------------------------

### Exercise 9. How Would You Decide the Classical (Frequentist) Way?

```{r frequentist}



```

------------------------------------------------------------------------

If you finished

## Post Lab Activity

A dealer hands you a **another coin** from the same set. You don't know
whether it is fair or biased. Your job is simple:

> **Is this coin biased?**

------------------------------------------------------------------------

### Your Data

Your data are in `lab3_post_data.csv`. There are 20 flips. You are going
to update posterior sequentially. After each flip, update the posterior
using the previous flip posterior as the new prior.

1.  Assign posterior values of the unbiased coin to `posterior1` at the
    first flip, `posterior2` at the second flip and `posterior3` at the
    third flip.
2.  Assign the last posterior value of the unbiased coin to
    `posterior20`
3.  Set `biased <- TRUE` if the coin is biased, `biased <- FALSE` if
    fair

> **AutoGrader requirements:**
>
> \- 4 scalars `posterior1` through `posterior3` and `posterior20` (they
> need to be numeric type in R)
>
> \- One logical `biased`: set `biased <- TRUE` if the coin is biased,
> `biased <- FALSE` if fair (logical type in R)

------------------------------------------------------------------------

```{r postlab activity}
# Please read !
# Due April 14, Tuesday 11:59pm
# Submit lab3.R to GradeScope
# You don't submit this Rmd file. 
# Don't setup your working directory. 
# Don't import libraries unless listed here.
# Please run your lab3.R file before you submit. 
# You should read the message from the autograder. Wait until you pass. 

library(dplyr)
library(ggplot2)
library(readr)


# Load post lab data
post_df  <- read.csv("lab3_post_data.csv")





```

------------------------------------------------------------------------
