---
title: "Lab 04: Conjugate Families — Beta-Binomial"
subtitle: ""
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    df_print: paged
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
```

------------------------------------------------------------------------

# The Coin Estimation Game

In Lab 3, we used Bayes' Rule with a **discrete** set of five candidate
coin types. Today we go further: we treat the true probability of heads
$\pi$ as a **continuous** unknown on $[0, 1]$, and use the
**Beta-Binomial conjugate family** to compute the posterior analytically

Your job is to estimate $\pi$ for your coin as precisely as possible and
report a credible interval for it.

Setup :

1\. Download all the csv files into working directory

2\. Set Rmd mode to `Visual` (from `Source` mode)

3.  Find your paired member

# Paired Lab

| Name | BUID |
|------|------|
|      |      |
|      |      |

------------------------------------------------------------------------

## Step 0.1: Choose Your Coin

```{r choose-coin}
data_types <- seq(1, 10)
your_study <- sample(data_types, 1)
cat("Your coin file: lab4_coin_", sprintf("%02d", your_study), ".csv",
    sep = "")
```

> **Note:** Load the file matching your assigned number, e.g.,
> `lab4_coin_03.csv`. Files run from `lab4_coin_01.csv` to
> `lab4_coin_10.csv`.

------------------------------------------------------------------------

## Step 0.2: Theory Review — The Beta-Binomial Model

### Why Conjugacy?

In Lab 3, we computed the posterior by multiplying the prior by the
likelihood for each discrete candidate and normalizing. With a
**continuous** $\pi \in [0,1]$, the normalizing integral
$\int_0^\infty f(\text{data} \mid \pi)\, f(\pi)\, d\pi$ is generally
hard to compute.

$$
f(\pi \mid \text{data}) = \frac{f(\text{data} \mid \pi)\, f(\pi)}
{\int_0^\infty f(\text{data} \mid \pi)\, f(\pi)\, d\pi}
$$

A **conjugate prior** is a special prior such that the posterior stays
in the **same distributional family** — making the update a simple
formula with no integration needed.

### The Beta Distribution — Prior for $\pi$

$$\pi \sim \text{Beta}(\alpha, \beta)$$

$$f(\pi) = \frac{\pi^{\alpha-1}(1-\pi)^{\beta-1}}{B(\alpha,\beta)},
\quad \pi \in [0,1]$$

where
$B(\alpha,\beta))= \frac{\Gamma(\alpha+\beta)}{\Gamma(\alpha) \Gamma(\beta)}$
with $$
\Gamma(Z) = \int_{0}^{\infty} x^{z-1} e^{-x} dx 
$$

Key facts:

$$E[\pi] = \frac{\alpha}{\alpha + \beta} \qquad
\text{Var}(\pi) =
\frac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta+1)}$$

The parameters $\alpha$ and $\beta$ act like **pseudo-observations**:

$\alpha-1$ : imaginary heads

$\beta-1$ : imaginary tails

$\alpha + \beta$ : how confident you were before the experiment

### The Binomial Likelihood

For $n$ flips with $x$ heads:

$$P(X = x \mid \pi) = \binom{n}{x} \pi^x (1-\pi)^{n-x}$$

The likelihood depends on the data only through the $x$

### The Beta-Binomial Posterior (Conjugate Update)

$$\pi \mid x \sim \text{Beta}(\alpha + x,\ \beta + n - x)$$

**Add** $x$ to $\alpha$, add $n - x$ to $\beta$. No integration, no
normalizing constant.

|   | Prior | Posterior |
|------------------------|:----------------------:|:----------------------:|
| Distribution | Beta$(\alpha, \beta)$ | Beta$(\alpha + x,\ \beta + n - x)$ |
| Mean | $\dfrac{\alpha}{\alpha+\beta}$ | $\dfrac{\alpha+x}{\alpha+\beta+n}$ |
| Pseudo-observations | $\alpha + \beta$ | $\alpha + \beta + n$ |

### Why? A Worked Example

Suppose you start with a **Beta(2, 2)** prior and observe $n = 40$ flips
with $x = 15$ heads (and 25 tails).

$$\pi \sim \text{Beta}(\alpha, \beta) = \text{Beta}(2,\ 2)$$

The prior mean is

$$E[\pi] = \frac{\alpha}{\alpha + \beta} = \frac{2}{2 + 2} = 0.50$$

This says: before seeing any data, we mildly believe the coin is fair,
with $\alpha + \beta = 4$ total pseudo-observations.

**The likelihood**

For $x = 15$ heads in $n = 40$ flips, the Binomial likelihood is

$$P(X = 15 \mid \pi) = \binom{40}{15}\, \pi^{15}\, (1 - \pi)^{25}$$

The key pieces are $\pi^{15}$ (15 heads) and $(1 - \pi)^{25}$ (25
tails). The binomial coefficient $\binom{40}{15}$ does not depend on
$\pi$, so it is just a constant in the update.

**The posterior construction: Multiply prior** $\times$ likelihood

$$f(\pi \mid x)  
  = \frac{ 
    \overbrace{\frac{\Gamma(\alpha+\beta)}{\Gamma(\alpha)\,\Gamma(\beta)} 
    \pi^{\alpha-1}(1-\pi)^{\beta-1}}^{\text{prior}}
  \times 
    \overbrace{\binom{n}{x} \pi^{x}(1-\pi)^{n-x}}^{\text{likelihood}} 
  }
  {\displaystyle\int_0^{1} 
    \frac{\Gamma(\alpha+\beta)}{\Gamma(\alpha)\,\Gamma(\beta)} 
    \pi^{\alpha-1}(1-\pi)^{\beta-1} 
    \cdot\binom{n}{x} \pi^{x}(1-\pi)^{n-x}\, d\pi}$$

Drop constants that do not depend on $\pi$ — they appear in both
numerator and denominator and cancel:

$$\propto \pi^{\alpha-1}(1-\pi)^{\beta-1} 
  \times \pi^{x}(1-\pi)^{n-x}$$

Collect the powers of $\pi$ and $(1-\pi)$:

$$\propto \pi^{(\alpha+x)-1}(1-\pi)^{(\beta+n-x)-1}$$

Plug in $\alpha = 2,\ \beta = 2,\ x = 15,\ n = 40$:

$$= \pi^{(2+15)-1}(1-\pi)^{(2+25)-1} 
  = \pi^{16}(1-\pi)^{26}$$

This is the **kernel** of a Beta distribution with updated parameters
$\alpha^* = \alpha + x$ and $\beta^* = \beta + n - x$. Since any valid
density must integrate to 1, the normalizing constant must be
$B(\alpha^*, \beta^*)$, giving the **exact closed-form posterior**:

$$\boxed{f(\pi \mid x) 
  = \frac{\pi^{(\alpha+x)-1}(1-\pi)^{(\beta+n-x)-1}}
         {B(\alpha+x,\ \beta+n-x)} 
  \sim \text{Beta}(\alpha + x,\ \beta + n - x)}$$

In our example:

$$f(\pi \mid x) 
  = \frac{\pi^{16}(1-\pi)^{26}}{B(17,\ 27)} 
  \sim \text{Beta}(17,\ 27)$$

> **Why does this work?** The Beta prior and Binomial likelihood share
> the same functional form in $\pi$ — both involve powers of $\pi$ and
> $(1 - \pi)$. Multiplying them together simply **adds the exponents**,
> producing another Beta kernel. This is exactly what it means for the
> Beta to be the **conjugate prior** for the Binomial likelihood: the
> posterior stays in the Beta family and the normalizing integral never
> needs to be computed explicitly.

**The posterior**

$$\pi \mid \text{data} \sim 
  \text{Beta}(\alpha + x,\ \beta + n - x) = 
  \text{Beta}(2 + 15,\ 2 + 25) = 
  \text{Beta}(17,\ 27)$$

No integration needed — the conjugate structure means the normalizing
constant works out automatically.

**Step 5: Compute posterior summaries**

$$\text{Posterior mean} 
  = \frac{\alpha + x}{\alpha + \beta + n} 
  = \frac{17}{17 + 27} 
  = \frac{17}{44} 
  \approx 0.386$$

$$\text{Sample proportion} = \frac{x}{n} = \frac{15}{40} = 0.375$$

$$\text{Prior mean} = 0.50$$

The posterior mean ($\approx 0.386$) sits **between** the prior mean
(0.50) and the sample proportion (0.375), pulled slightly toward 0.50
because the prior contributes $\alpha + \beta = 4$ pseudo-observations
out of $\alpha + \beta + n = 44$ total. As $n$ grows, the posterior mean
converges to the sample proportion.

**Step 6: Verify in R**

```{r worked-example}
# known values
alpha_0 <- 2
beta_0  <- 2
x       <- 15
n       <- 40

# conjugate update
alpha_new <- alpha_0 + x
beta_new  <- beta_0  + (n - x)

# summaries
prior_mean <- alpha_0  / (alpha_0  + beta_0)
post_mean  <- alpha_new / (alpha_new + beta_new)
samp_prop  <- x / n

cat("Prior:             Beta(", alpha_0,   ",", beta_0,    ")\n")
cat("Posterior:         Beta(", alpha_new, ",", beta_new,  ")\n")
cat("Prior mean:        ",      round(prior_mean, 4),      "\n")
cat("Sample proportion: ",      round(samp_prop,  4),      "\n")
cat("Posterior mean:    ",      round(post_mean,  4),      "\n")
cat("Prior weight:      ",
    round((alpha_0 + beta_0) / (alpha_0 + beta_0 + n) * 100, 1),
    "%\n")
```

```{r worked-example-plot, fig.height=4, fig.width=7}
pi_grid <- seq(0, 1, length.out = 500)

tibble(
  pi   = rep(pi_grid, 2),
  dens = c(dbeta(pi_grid, alpha_0,   beta_0),
           dbeta(pi_grid, alpha_new, beta_new)),
  dist = rep(
    c(paste0("Prior: Beta(",      alpha_0,   ", ", beta_0,   ")"),
      paste0("Posterior: Beta(",  alpha_new, ", ", beta_new, ")")),
    each = 500)
) |>
  mutate(dist = factor(dist, levels = unique(dist))) |>
  ggplot(aes(x = pi, y = dens, color = dist, fill = dist)) +
  geom_line(linewidth = 1.2) +
  geom_area(alpha = 0.10, position = "identity") +
  geom_vline(xintercept = prior_mean, linetype = "dashed",
             color = "steelblue", linewidth = 0.8) +
  geom_vline(xintercept = samp_prop,  linetype = "dashed",
             color = "tomato",    linewidth = 0.8) +
  geom_vline(xintercept = post_mean,  linetype = "solid",
             color = "darkgreen", linewidth = 0.8) +
  annotate("text", x = prior_mean - 0.01, y = 6.5,
           label = "Prior mean\n0.50",   color = "steelblue",
           hjust = 1, size = 3.2) +
  annotate("text", x = samp_prop  - 0.01, y = 5.5,
           label = "Sample prop\n0.375", color = "tomato",
           hjust = 1, size = 3.2) +
  annotate("text", x = post_mean  + 0.01, y = 6.0,
           label = "Post mean\n0.386",   color = "darkgreen",
           hjust = 0, size = 3.2) +
  scale_color_manual(values = c("steelblue", "tomato")) +
  scale_fill_manual( values = c("steelblue", "tomato")) +
  labs(title    = "Prior vs. Posterior — Worked Example",
       subtitle = "n = 40 flips  |  x = 15 heads",
       x = expression(pi ~ "(P(Heads))"),
       y = "Density", color = NULL, fill = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```

**Summary table**

|                     |   Prior    |   Posterior   |
|---------------------|:----------:|:-------------:|
| Distribution        | Beta(2, 2) | Beta(17, 27)  |
| $\alpha$            |     2      | 17 $= 2 + 15$ |
| $\beta$             |     2      | 27 $= 2 + 25$ |
| Mean                |   0.500    |     0.386     |
| Pseudo-observations |     4      | 44 $= 4 + 40$ |

> **Key intuition:** Adding $x = 15$ to $\alpha$ records the observed
> heads; adding $n - x = 25$ to $\beta$ records the observed tails. The
> total pseudo-observation count grows from 4 to 44, reflecting the
> information gained from 40 real flips. The prior's influence shrinks
> from $4/44 \approx 9\%$ of the total weight — nearly negligible — and
> the posterior mean is pulled strongly toward the sample proportion.

------------------------------------------------------------------------

## Step 0.3: Load Data

```{r load-data}
my_coin_file <- "lab4_coin_01.csv"   # TODO: change to your assigned file
coin_df      <- read.csv(my_coin_file)
head(coin_df, 10)
```

```{r data-summary}
n_flips     <- nrow(coin_df) 
n_heads     <- sum(coin_df$result)
n_tails     <- n_flips - n_heads
sample_prop <- round(n_heads / n_flips, 3)

cat("File:             ", my_coin_file, "\n")
cat("Total flips:      ", n_flips,      "\n")
cat("Heads:            ", n_heads,      "\n")
cat("Tails:            ", n_tails,      "\n")
cat("Sample proportion:", sample_prop,  "\n")
```

------------------------------------------------------------------------

## Step 0.4: Setting Up the Notation

The unknown is the **continuous** probability of heads:

$$\pi \in [0, 1]$$

Unlike Lab 3, $\pi$ is not restricted to five values — it can be
anything in $[0,1]$. We model uncertainty about $\pi$ with a **Beta
distribution** as the prior.

------------------------------------------------------------------------

## Step 1: First Approach — Using All 50 Flips at Once

### Exercise 1. Choose Your Prior

We use a **Beta** prior —

```{r prior-setup}
# Decide your initial alpha and beta values
alpha_prior <- # TODO
beta_prior  <- # TODO

prior_mean <- alpha_prior / (alpha_prior + beta_prior)
prior_var  <- (alpha_prior * beta_prior) /
              ((alpha_prior + beta_prior)^2 * (alpha_prior + beta_prior + 1))
prior_sd   <- sqrt(prior_var)

cat("Prior: Beta(", alpha_prior, ",", beta_prior, ")\n")
cat("Prior mean:      ", round(prior_mean, 4), "\n")
cat("Prior SD:        ", round(prior_sd,   4), "\n")
cat("Pseudo-obs total:", alpha_prior + beta_prior, "\n")
```

```{r prior-plot, fig.height=3.5, fig.width=6}
pi_grid <- seq(0, 1, length.out = 500)

tibble(pi      = pi_grid,
       density = dbeta(pi_grid, alpha_prior, beta_prior)) |>
  ggplot(aes(x = pi, y = density)) +
  geom_line(color = "steelblue", linewidth = 1.2) +
  geom_area(fill = "steelblue", alpha = 0.15) +
  geom_vline(xintercept = 0.5, linetype = "dashed",
             color = "gray50", linewidth = 0.8) +
  labs(title    = paste0("Prior: Beta(", alpha_prior, ", ", beta_prior, ")"),
       subtitle = paste0("Mean = ", round(prior_mean, 3),
                         "  |  SD = ", round(prior_sd, 3),
                         "  |  Pseudo-obs = ", alpha_prior + beta_prior),
       x = expression(pi ~ "(P(Heads))"), y = "Density") +
  theme_minimal(base_size = 13)
```

> **Exercise 1.2 Why did you choose the values for** $\alpha$ and
> $\beta$ **?**
>
> answer :

------------------------------------------------------------------------

### Exercise 2. Compute the Posterior (Using the full data)

Using the conjugate update rule:
$\alpha' = \alpha + x \qquad \beta' = \beta + (n - x)$

```{r posterior-batch}
alpha_post <- #TODO: alpha for posterior 
beta_post  <- #TODO: beta for posterior 

post_mean <- #TODO
post_var  <- #TODO
post_sd   <- sqrt(post_var)
post_mode <- (alpha_post - 1) / (alpha_post + beta_post - 2)

cat("Posterior: Beta(", alpha_post, ",", beta_post, ")\n")
cat("Posterior mean:   ", round(post_mean, 4), "\n")
cat("Posterior mode:   ", round(post_mode, 4), "\n")
cat("Posterior SD:     ", round(post_sd,   4), "\n")
cat("Sample proportion:", sample_prop,          "\n")
```

------------------------------------------------------------------------

### Exercise 3. Visualize Prior vs. Posterior

```{r prior-posterior-plot, fig.height=4, fig.width=7}
tibble(
  pi   = rep(pi_grid, 2),
  dens = c(dbeta(pi_grid, alpha_prior, beta_prior),
           dbeta(pi_grid, alpha_post,  beta_post)),
  dist = rep(
    c(paste0("Prior: Beta(", alpha_prior, ", ", beta_prior, ")"),
      paste0("Posterior: Beta(", alpha_post, ", ", beta_post, ")")),
    each = 500)
) |>
  mutate(dist = factor(dist, levels = unique(dist))) |>
  ggplot(aes(x = pi, y = dens, color = dist, fill = dist)) +
  geom_line(linewidth = 1.2) +
  geom_area(alpha = 0.10, position = "identity") +
  geom_vline(xintercept = sample_prop, linetype = "dashed",
             color = "gray40", linewidth = 0.9) +
  annotate("text", x = sample_prop + 0.02,
           y = max(dbeta(pi_grid, alpha_post, beta_post)) * 0.25,
           label = paste0("x/n = ", sample_prop),
           color = "gray40", hjust = 0, size = 3.5) +
  scale_color_manual(values = c("steelblue", "tomato")) +
  scale_fill_manual(values  = c("steelblue", "tomato")) +
  labs(title    = paste0("Prior vs. Posterior — ", my_coin_file),
       subtitle = paste0(n_heads, " heads out of ", n_flips, " flips"),
       x = expression(pi ~ "(P(Heads))"),
       y = "Density", color = NULL, fill = NULL) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```

**Exercise 3.2 :** Describe how the posterior differs from the prior.
Why does the posterior concentrate around the sample proportion?

**Answer:**

------------------------------------------------------------------------

## Step 2: Sequential Updating — (Flip by Flip)

Instead of using all 50 flips at once, we can update the Beta
distribution **one flip at a time**:

-   Observing one trial at a time with success $s \in {0,1}$

-   Prior : Beta(alpha, beta)

-   Posterior : Beta(alpha+s, beta+(1-s))

    Updates $\alpha → \alpha + s$ and $\beta → \beta + (1 − s)$

### Step 2.1: Manual Table — First 3 Flips

```{r manual-3-flips 0}
flips <- coin_df$result   # 1 = Heads, 0 = Tails

cat("First 3 flips:", flips[1], flips[2], flips[3],
    "(1=H, 0=T)\n\n")

# --- Flip 0: Prior ---
a1 <- #TODO
b1 <- #TODO
cat("alpha0 at flip 0:", a1, "\n")
cat("beta0 at flip 0:", b1)
```

```{r manual-3-flips 1}
# --- Flip 1 ---
a2 <- #TODO
b2 <-  #TODO
cat("alpha0 at flip 1:", a2, "\n")
cat("beta0 at flip 1:", b2)
```

```{r manual-3-flips 2}
# --- Flip 2 ---
a3 <-  #TODO
b3 <-  #TODO
cat("alpha0 at flip 2:", a3, "\n")
cat("beta0 at flip 2:", b3)
```

```{r manual-3-flips 3}
# --- Flip 3 ---
a4 <-  #TODO
b4 <-  #TODO
cat("alpha0 at flip 3:", a4, "\n")
cat("beta0 at flip 3:", b4)
```

```{r manual-3-flips}
flips <- coin_df$result

cat("First 3 flips:", flips[1], flips[2], flips[3], "(1=H, 0=T)\n\n")

tibble(
  Stage       = c("Prior", "After flip 1", "After flip 2", "After flip 3"),
  Flip_result = c(NA, flips[1], flips[2], flips[3]),
  alpha       = c(a1, a2, a3, a4),
  beta        = c(b1, b2, b3, b4),
  Mean        = round(c(a1,a2,a3,a4) /
                      (c(a1,a2,a3,a4) + c(b1,b2,b3,b4)), 4),
  SD          = round(sqrt(
    c(a1,a2,a3,a4) * c(b1,b2,b3,b4) /
    ((c(a1,a2,a3,a4)+c(b1,b2,b3,b4))^2 *
     (c(a1,a2,a3,a4)+c(b1,b2,b3,b4)+1))
  ), 4)
)
```

------------------------------------------------------------------------

### Step 2.2: Full Sequential Updating (using the logic above)

```{r sequential-update}
n <- n_flips
a_seq <- rep(0, n + 1)
b_seq <- rep(0, n + 1)
a_seq[1] <- alpha_prior
b_seq[1] <- beta_prior

for (i in 1:n) {
  a_seq[i + 1] <-  #TODO
  b_seq[i + 1] <-  #TODO
}

post_mean_seq <- a_seq / (a_seq + b_seq)
post_sd_seq   <- sqrt((a_seq * b_seq) /
                      ((a_seq + b_seq)^2 * (a_seq + b_seq + 1)))

cat("Sequential updating complete.\n")
cat("Final posterior: Beta(",
    tail(a_seq,1), ",", tail(b_seq,1), ")\n")
cat("Final mean:", round(tail(post_mean_seq,1), 4),
    "| Final SD:", round(tail(post_sd_seq,1), 4), "\n")
```

```{r snapshot-table}
snapshots <- c(0, 5, 10, 20, 50)

map_dfr(snapshots, function(k) {
  tibble(
    `After flips` = k,
    alpha         = a_seq[k + 1],
    beta          = b_seq[k + 1],
    Mean          = round(post_mean_seq[k + 1], 4),
    SD            = round(post_sd_seq[k + 1],   4)
  )
})
```

```{r trajectory-plot, fig.height=4, fig.width=8}
tibble(flip = 0:n,
       mean = post_mean_seq,
       lo   = post_mean_seq - 2 * post_sd_seq,
       hi   = post_mean_seq + 2 * post_sd_seq) |>
  ggplot(aes(x = flip, y = mean)) +
  geom_ribbon(aes(ymin = lo, ymax = hi),
              fill = "steelblue", alpha = 0.2) +
  geom_line(color = "steelblue", linewidth = 1.1) +
  geom_point(data = ~ filter(., flip %in% snapshots),
             color = "tomato", size = 3) +
  geom_hline(yintercept = 0.5, linetype = "dashed",
             color = "gray50", linewidth = 0.8) +
  annotate("text", x = 43, y = 0.52,
           label = "Fair coin (0.50)", color = "gray50", size = 3.5) +
  labs(title    = "Posterior Mean ± 2 SD After Each Flip",
       subtitle = "Shaded band = ±2 posterior SDs  |  Red dots = snapshots",
       x = "Number of flips",
       y = expression("Posterior mean of " ~ pi)) +
  ylim(0, 1) +
  theme_minimal(base_size = 13)
```

```{r posterior-snapshots-plot, fig.height=5, fig.width=10}
map_dfr(snapshots, function(k) {
  tibble(
    pi    = pi_grid,
    dens  = dbeta(pi_grid, a_seq[k+1], b_seq[k+1]),
    label = paste0("After ", k, " flips — Beta(",
                   a_seq[k+1], ",", b_seq[k+1], ")")
  )
}) |>
  mutate(label = factor(label, levels = unique(label))) |>
  ggplot(aes(x = pi, y = dens, color = label)) +
  geom_line(linewidth = 1.1) +
  scale_color_brewer(palette = "RdYlBu", direction = -1) +
  labs(title  = "Posterior Beta Distribution at Each Snapshot",
       x      = expression(pi ~ "(P(Heads))"),
       y      = "Density", color = NULL) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom")
```

------------------------------------------------------------------------

## Exercise 4. Are the Posteriors the Same?

Compare the batch update (all 50 flips at once) to the sequential update
(flip by flip). Do they agree?

**Answer:**

## Exercise 5. Credible Interval

Obtain A **95% credible interval** $[L, U]$ satisfies
$P(L \leq \pi \leq U \mid \text{data}) = 0.95$.

Obtain A **90% credible interval** $[L, U]$ satisfies
$P(L \leq \pi \leq U \mid \text{data}) = 0.90$. Unlike a frequentist
confidence interval, this is a direct probability statement about $\pi$.

```{r credible-interval}
ci_95 <- #TODO: get 95% credible interval for pi, hint;use qbeta()
ci_90 <- #TODO: get 90% credible interval for pi, hint;type ?qbeta

cat("Posterior: Beta(", alpha_post, ",", beta_post, ")\n")
cat("95% Credible Interval: [",
    round(ci_95[1], 4), ",", round(ci_95[2], 4), "]\n")
cat("90% Credible Interval: [",
    round(ci_90[1], 4), ",", round(ci_90[2], 4), "]\n")
```

```{r ci-plot, fig.height=4, fig.width=7}
shade_df <- tibble(pi   = pi_grid,
                   dens = dbeta(pi_grid, alpha_post, beta_post)) |>
  filter(pi >= ci_95[1], pi <= ci_95[2])

peak_dens <- max(dbeta(pi_grid, alpha_post, beta_post))

tibble(pi = pi_grid, dens = dbeta(pi_grid, alpha_post, beta_post)) |>
  ggplot(aes(x = pi, y = dens)) +
  geom_line(color = "steelblue", linewidth = 1.2) +
  geom_area(data = shade_df, fill = "steelblue", alpha = 0.35) +
  geom_vline(xintercept = ci_95, linetype = "dashed",
             color = "tomato", linewidth = 0.9) +
  annotate("text", x = ci_95[1] - 0.01, y = peak_dens * 0.55,
           label = round(ci_95[1], 3), color = "tomato",
           hjust = 1, size = 3.5) +
  annotate("text", x = ci_95[2] + 0.01, y = peak_dens * 0.55,
           label = round(ci_95[2], 3), color = "tomato",
           hjust = 0, size = 3.5) +
  labs(title    = paste0("Posterior Beta(", alpha_post, ", ", beta_post,
                          ") with 95% Credible Interval"),
       subtitle = "Shaded region contains 95% of posterior probability",
       x = expression(pi ~ "(P(Heads))"), y = "Density") +
  theme_minimal(base_size = 13)
```

## Exercise 5.2. Interpret the credible intervals.

**Answer:**

------------------------------------------------------------------------

## Exercise 6. What Is Your Final Estimate and what do you see?

**Answer:**

------------------------------------------------------------------------

## Exercise 7. How Would the Frequentist Approach This?

```{r frequentist}
p_hat      <- n_heads / n_flips
se         <- sqrt(p_hat * (1 - p_hat) / n_flips)
ci_freq_95 <- p_hat + c(-1, 1) * qnorm(0.975) * se

cat("Sample proportion:       ", round(p_hat, 4),       "\n")
cat("Standard error:          ", round(se,    4),        "\n")
cat("95% Confidence Interval: [",
    round(ci_freq_95[1], 4), ",", round(ci_freq_95[2], 4), "]\n")



```

------------------------------------------------------------------------

## Post Lab Activity

A casino is testing a new roulette-style coin for a promotional game.
They claim the coin is **fair** ($\pi = 0.50$). A player suspects it may
be biased toward heads.

Your data are in `lab4a_post_data.csv` — 60 flips of this coin.

Update a Beta posterior **in three batches of 20 flips** and report the
posterior after each batch, then make a final decision.

> **AutoGrader requirements:**
>
> -   `post_alpha1`, `post_beta1` — posterior parameters after batch 1
>     (flips 1–20)
> -   `post_alpha2`, `post_beta2` — posterior parameters after batch 2
>     (flips 1–40)
> -   `post_alpha3`, `post_beta3` — posterior parameters after batch 3
>     (all 60 flips)
> -   `ci_lower`, `ci_upper` — bounds of 95% credible interval after all
>     60 flips
> -   `biased` — logical: `TRUE` if you conclude the coin is biased?

------------------------------------------------------------------------

```{r postlab activity}
# Please read !
# Submit lab4.R to GradeScope
# You don't submit this Rmd file. 
# Don't setup your working directory. 
# Don't import libraries unless listed here.
# Please run your lab4.R file before you submit. 
# You should read the message from the autograder. Wait until you pass. 
# Pass means 5/5 

library(dplyr)
library(ggplot2)
library(readr)

post_df    <- read.csv("lab4_post_data.csv")
post_flips <- post_df$result
n_post     <- length(post_flips)

a0_post <- 2
b0_post <- 2

cat("Starting prior: Beta(", a0_post, ",", b0_post, ")\n")
cat("Prior mean:", a0_post / (a0_post + b0_post), "\n")





```
